### 2019-08-06 paper/paper.tex
included zeroth draft of paper texfile and bibtex file

% Whisker barrel generation paper equations, for pasting into Google
% docs draft.

% While writing, don't stop for errors
%\nonstopmode

\documentclass[a4paper,11pt]{article}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{enumerate}
\usepackage{times}
\usepackage{color}
\usepackage{parskip}
\usepackage{afterpage}
\usepackage{setspace}
\usepackage[modulo]{lineno}
%\usepackage{xltxtra}
%\usepackage{xunicode}
\usepackage{gensymb}

\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{comment}
\usepackage{textcomp}
\usepackage{natbib}

\usepackage[labelfont=bf,labelsep=period,justification=justified]{caption}

\doublespacing
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm
\textheight 21cm
\makeatletter
\makeatother
\date{}
\pagestyle{plain}

%\newcommand{\dvrg}{\nabla\vcdot\nabla}
\newcommand{\mb}[1]{\mathbf{#1}}
\makeatletter
\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
\makeatother

%\bibliographystyle{plos2015}
\bibliographystyle{apalike}

\begin{document}
%\begin{center}
\begin{flushleft}
{\huge
\textbf{Self-organization of whisker barrels in simulation requires competition between thalamocortical axons}
 \vspace{5em}
}
%\end{center}

Sebastian~S.~James$^{1*}$, Stuart~P.~Wilson$^{1}$ \vspace{1em}\\
$^{1}$Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.\vspace{1em}\\
$^{*}$Corresponding author email: Seb.James@Sheffield.ac.uk\vspace{1em}\\
Short title: Self-organising whisker barrels\vspace{5em}\\
\end{flushleft}

\newpage{}
%\raggedright

\section*{Abstract}

Understanding how genetic information and spontaneous pattern formation together give rise to the whisker barrels is an important challenge for developmental neurobiology. The barrel patterning that coalesces in layer IV has been described formally as a Voronoi tesselation, which is indicative of an underlying developmental process based on local competitive interactions. If pattern formation relies only on such intrinsic self-organising processes, how can the somatotopic order at the periphery be maintained in the projections from brainstem to thalamus and from thalamus to cortex? To investigate, we began by extending a computational model of neocortical arealization by Karbowski \& Ermentrout (2004; J. Comp. Neurosci.) to generate fields on a two-dimensional cortical sheet. Central to this model is an assumption that axonal branching and synaptogenesis are coupled, and that thalamic projections compete for cortical territory under the influence of molecular signalling gradients. We found that these processes alone are incapable of generating patterns of cortical fields that are more intricate than the patterns already present in molecular guidance signals. However, a full barrel-field pattern emerges under the guidance of as few as four signalling gradients when the model is extended to introduce competition between the branches of axons that originate from different thalamic barreloids. We show that the key requirements for the emergence of realistic (Voronoi tesselation) patterning are i) at each cortical location thalamocortical projections compete for a limited number of available connections, ii) at each location the axonal branching rate of each thalamocortical projection is reduced by the total branching density of the other projections, and iii) the density of axon branching for each thalamocortical projection is conserved. We demonstrate how a purely self-organising system can initialise a pattern which can then be faithfully replicated in multiple neural structures. We show that this minimal model of barrel self-organisation provides a parsimonious account of data from a range of classic studies in which intrinsic and extrinsic factors have been experimentally manipulated.

\section*{Introduction}

\modulolinenumbers{}
\linenumbers

Spatial patterns in neural connectivity provide clues about the constraints under which brains evolve and develop. Whether or not patterning governs information processing, models of the development of information processing circuits can be validated in terms of their ability to account for biological patterning (\citealp{Purves1992}; see also \citealp{Wilson2015,Bednar2016}). One of the most distinctive patterns of neural connectivity is found in the rodent barrel cortex (\citealp{Woolsey1948,Woolsey1970,Welker1974}, see \citealp{Fox2008}). Understanding how this particular pattern forms might shed light on the interplay of genetic and extrinsic factors that shape cortical circuits, and on a wealth of anatomical, physiological, and behavioural data obtained using this model system.

The barrels are large cortical columns, identifiable from birth in layer 4 of rodent primary somatosensory cortex as dense clusters of thalamocortical axons, which are enclosed by borders a few neurons thick from postnatal day 3 \citep{Erzurumlu2012}. In the plane tangential to the cortical surface the barrels constitute a map of the arrangement of the whiskers on the face, such that the cells of adjacent barrels respond most strongly and most quickly to deflection of adjacent whiskers \citep{Armstrong-James1992}. The barrels reflect subcortical whisker maps comprising cellular aggregates known as barrelettes in the brainstem and barreloids in the thalamus \citep{Killackey1980}.

Each whisker and its corresponding barrelette, barreloid, and barrel can be identified by its position on a two-dimensional grid \citep{Zucker1969,Killackey1980,Loos1976}. The barrels are normally up to twice as long in the axis aligned to the grid ‘rows’ compared to the ‘arcs’, and the rows are visible from day 2 \citep{Rebsam2002} whereas the arcs become distinct by day 3 \citep{Erzurumlu2012,Rebsam2002}. In metabotropic glutamate receptor 5  (mGlu5) knockouts the barrel rows develop but the arcs do not (\citealp{Hannan2001} see also \citealp{Fox2008}), thus distinct processes likely guide development of the barrels with respect to the orthogonal grid axes.

Barrel formation requires afferent input, driven by whisker stimulation and by thalamic calcium waves that propagate via mGlu5-receptor activation (\citealp{Anton2019}, see also \citealp{Anton2018}). Patterning also depends on a complex network of axon guidance processes, which includes interactions between guidance molecules like ephrin-A5 and A7 \citep{Miller2006} and between adhesion molecules like cadherin-6 and 8 \citep{Greig2013,Bishop2000}, and the binding of p75 to neurotropins (\citealp{Shimogori2005,Bishop2002}; see also \citealp{Dye2011a,Dye2011b}). This network is orchestrated by interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2, Pax6, Sp8, and Coup-tf1 \citep{Shimogori2005,Bishop2000}, which are expressed in gradients that mark orthogonal axes and can be manipulated to cause the barrels to stretch, shrink, shift, skew, and even duplicate \citep{Shimogori2005,Assimacopoulos2012,Borello2014,Sahara2007,Ypsilanti2016,Sur2005}.

The boundaries of the barrels form polygonal shapes that together create a Voronoi tesselation across the cortical sheet, wherein all points enclosed by a given boundary are closer to its center than to the center of any other boundary \citep{Senft1991}. This patterning suggests that the barrel topology may be preserved in the initial projection of the thalamocortical axons into the cortex, and that a barrel forms by the lateral branching of the axons from their initial centres of mass, which ceases upon contact with the axons branching from adjacent centres. According to this model the barrel pattern could be entirely determined by the distribution of the initial centre points, which would itself be determined by guidance cues. However, the assumption of pre-arranged centres is difficult to resolve with the observation that thalamocortical axons arrive in the cortical plate as an undifferentiated bundle prior to barreloid formation (see \citealp{Erzurumlu2012} for a review).

Alternatively, reaction-diffusion dynamics could generate a Voronoi-like tesselation in the cortex without the need for pre-arranged centres, by amplifying characteristic modes in the noisy initial distribution of bundled axon branches, as a net effect of short-range cooperative and longer-range competitive interactions between cells. According to this model the barrel pattern would be determined by the relative strength of these interactions and by the shape of the cortical field boundary. However, reaction-diffusion dynamics alone are unlikely to be able to account for the precise co-registration and topographic connectivity between emergent domains in thalamus and cortex, the non-uniform sizes and spacing of the barrels to form rows versus arcs, and the contribution of gene expression gradients.

As theories of barrel patterning, the centre-point and reaction-diffusion models are not mutually exclusive. Pre-organised centres in the initial conditions could bias reaction-diffusion processes to generate the specific somatotopic arrangement of barrels more reliably, and similarly the mechanisms by which axons branch laterally from barrel centres may involve a tension between cooperative and competitive interactions. However, a proof that barrel patterning can in principle emerge within an undifferentiated bundle of thalamocortical axons, based only on local interactions between cells, would remove the need to invoke a global supervisor, i.e., to assume a separate stage and/or extrinsic mechanism for pre-organising thalamocortical connections.

To this end, the aim of the present study is to establish whether it is possible for the observed arrangement of barrels to self-organise in a system governed by reaction-diffusion dynamics, under the guidance of molecular signalling gradients, and in the absence of pre-defined centre-points. Our approach is to represent in a computational model the minimal set of biological assumptions that might lead to the emergence of barrel patterning in simulation, and, if necessary, to add to this minimal set only assumptions that can be represented by the local exchange of information between adjacent cells. The emergence of barrel patterning under such conditions would serve as an existence proof of the plausibility of the hypothesis that barrel development is a self-organising process.

\section*{Models}

\cite{Karbowski2004} developed what remains, to the best of our knowledge, the only dynamical model of cortical arealization capable of generating predictions about the influence of signalling molecules on the size, shape, and topological organization of emergent cortical fields. This model describes competition between thalamocortical projections for territory along a one-dimensional string of cortical cells, e.g., along the anterior-posterior axis. The equations of the model define how the density of connections, $c$, and the density of axon branches, $a$, interact at time $t$ at each cortical location $x$, for $N$ thalamocortical projections indexed by $i$. The tesselation of the barrels that we investigate is a two-dimensional structure, hence we first extended the original model for simulation in two spatial dimensions, and describe it here as such (i.e., suibstituting $x$ for $\mb{x}$ and $\frac{d}{dx}$ for $\nabla$).

The model was derived from the assumption that the rates of synaptogenesis and thalamocortical axon branching are reciprocally coupled, resulting in an expression for the change in $c_i$ as a function of $a_i$, and for the change in $a_i$ as a function of $c_i$. Synaptogenesis is modelled  as

\begin{equation} \label{eq:Karb2DExt_dc}
\frac{\partial c_i(\mb{x},t)}{\partial t} =-\alpha c_i(\mb{x},t) +\beta  \left(1 - \sum_{j=1}^{N} c_j(\mb{x}, t)\right).[a_i(\mb{x},t)]^k.
\end{equation}

Accordingly when the final term evaluates to zero, the connection density decays exponentially at a rate given by the decay constant $\alpha$. This occurs, at a given location, when the sum of connections is one. Otherwise, the connection density is assumed to vary non-linearly ($k>1$) with changes in the axon branching density.

The change in density of axon branches is assumed to depend also on the influence of guidance molecules and on the density of branches at adjacent cortical locations:

%
\begin{equation} \label{eq:Karb2DExt_da}
\frac{\partial a_i(\mb{x},t)}{\partial t} = \nabla\vcdot\left(D \nabla a_i(\mb{x},t)-a_i(\mb{x},t)\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i(\mb{x},t)}{\partial t} + \chi
\end{equation}

The first term on the right describes the spatial gradient (indicated by $\nabla$) of a quantity referred to in models of heat diffusion as the `flux' (enclosed by parentheses). The central idea in such models is that the change in temperature (over time) at a given point on a surface is proportional to the gradient in the temperature (over space) at that location. Substituting heat for the number of branches, axons from a given thalamic projection are thus assumed to branch laterally across the cortical sheet at a rate given by the diffusion constant $D$.

Lateral branching is augmented by a measure of the local alignment between the density of branches and the gradient in the concentration of each of $M$ distinct molecular signals, $\rho(\mb{x})$. The influence of a given signal (indexed by $j$) on a given projection (indexed by $i$), is determined by the parameter $\gamma_i,j$. Interactions for which this value is positive encourage axon branching in the direction in which the concentration of the signal is increasing, whereas negative interactions encourage branching in the opposite direction. Hence the axons of a given projection are assumed to branch towards either higher or lower concentrations of each signalling molecule.

The second term on the right of equation 2 represents an assumption that the number of axon branches decreases when the number of connections being made increases, and vice versa, i.e., that branches are required to form connections. Note that we introduce the term $\chi=0$ here, so that we might subsequently redefine $\chi$ in order to introduce the competition between thalamic projections that we found to be necessary for the emergence of a Voronoi tesselation.

Two points are important to emphasise here. First, notice that the only difference between thalamic projections is the $M$ interaction strengths ($\gamma$) that are chosen to define each projection. Hence any (reliable) differences between the size, shape, and location of simulated cortical fields are due to differences in the choice of these values only. Second, the formulation of the model is at this point entirely local, in the sense that simulating the dynamics does not require any information to be communicated from one cortical cell to any but its immediate neighbours.

In addition to the choice of initial conditions (for the fields $c(\mb{x},t=0)$ and $a(\mb{x},t=0)$) and the axon guidance fields $\rho(\mb{x})$, the free parameters of the model are the number of thalamocortical projections ($N$), constants $\alpha$, $\beta$, $k$, and $D$, and the $N\times M$ interaction strengths, $\gamma$.

Equations 1 and 2 were solved on a hexagonal lattice of grid elements represent the cortical sheet ($\mb{x}$), under the assumption that axon branches cannot extend beyond the boundary of the tissue, i.e., $\mb{J}_i(\mb{x},t) \bigg\rvert_{boundary} = 0$. We developed software tools for imposing arbitrary boundary shapes, specified in a scalable vector graphics (svg) file format, on the lattice. The dynamics were integrated using fourth-order Runge-Kutta for ordinary differential equation 1, and using a {\color{red} finite difference method} for the flux term (see Supplementary Materials).

\section*{Results}

%\begin{equation}
%\chi=\frac{\epsilon_i  a_i}{N-1} \sum_{j \ne i}^{N} a_j(\mb{x}, t)^l
%\end{equation}

%And normalization step

%\begin{equation}\label{eq:norm_comp8}
%a_i' = n_e \; \frac {a_i} {\textstyle \sum_j^{n_e} a_{i,j} }
%\end{equation}

%($a_i'$ becomes $a_i$ in the next computational step.)

\section*{Discussion}

We set out to test the hypothesis that barrel patterning may emerge via self-organising processes. We first extended an existing model of neocortical arealization to simulate pattern formation in two dimensions, inheriting from this model the assumption that axonal branching and synaptogenesis are coupled, and that thalamic projections compete for cortical territory under the influence of molecular signalling gradients. Realistic barrel patterning emerged only when terms were added to represent the assumption that  thalamocortical projections compete for a limited supply of connections with cortical cells, that their branching rates are reduced by the branching of other thalamocortical projections, and that the density of axon branching is conserved for each projection. The net effect of these local interactions is a Voronoi tesselation in which localised clusters of thalamic projections become somatotopically organised on the cortical sheet according to the relative values of coefficients specifying the rates at which axons are attracted and repelled by orthogonal pairs of molecular signal gradients. According to this model, the pattern of barrels is determined by the competitive and cooperative interactions between cells relative to the shape of the cortical field boundary. In simulation, we systematically investigated the origin of the asymmetry between the barrel rows and arcs, and identified {\color{red} XXXXX} as the most likely candidate for the elongation of barrel shapes along the rows.

\bibliography{bibfile}

\end{document}

%%%%% ORIGINAL EQUATIONS MAINTAINED AS NOTES HERE FOR REFERENCE

\section{Karbowski system with N TCs and M gradients}
%
\begin{equation} \label{eq:Karb2D_dc}
\frac{\partial c_i(\mb{x},t)}{\partial t} = -\alpha c_i(\mb{x},t) + \beta n(\mb{x},t)
[a_i(\mb{x},t)]^k
\end{equation}
%
\begin{equation} \label{eq:Karb2D_conserve}
n(\mb{x},t) = 1 - \sum_{i=1}^{N} c_i(\mb{x}, t)
\end{equation}
%
\begin{equation} \label{eq:Karb2D_da}
\frac{\partial a_i(\mb{x},t)}{\partial t}
= \nabla\vcdot\mb{J}_i(\mb{x},t) - \frac{\partial c_i(\mb{x},t)}{\partial t}
\end{equation}

with the flux current

\begin{equation} \label{eq:Karb2D_J}
\mb{J}_i(\mb{x},t) = D \nabla a_i(\mb{x},t) - a_i
\sum_{j=1}^M \big(\gamma_{i,j} \nabla\rho_j(\mb{x}) \big)
\end{equation}

The boundary condition applied to these equations is:

\begin{equation}
\mb{J}_i(\mb{x},t) \bigg\rvert_{boundary} = 0
\end{equation}

\section{Karbowski system with N TCs and M gradients plus competition}

%
\begin{equation} \label{eq:Karb2DExt_dc}
\frac{\partial c_i(\mb{x},t)}{\partial t} = -\alpha c_i(\mb{x},t) + \beta n(\mb{x},t)
[a_i(\mb{x},t)]^k
\end{equation}
%
\begin{equation} \label{eq:Karb2DExt_conserve}
n(\mb{x},t) = 1 - \sum_{i=1}^{N} c_i(\mb{x}, t)
\end{equation}
%
\begin{equation} \label{eq:Karb2DExt_da}
\frac{\partial a_i(\mb{x},t)}{\partial t}
= \nabla\vcdot\mb{J}_i(\mb{x},t) - \frac{\partial c_i(\mb{x},t)}{\partial t}- \frac{\epsilon_i  a_i}{N-1} \sum_{j \ne i}^{N} a_j(\mb{x}, t)^l
\end{equation}

And normalization step

\begin{equation}\label{eq:norm_comp8}
a_i' = n_e \; \frac {a_i} {\textstyle \sum_j^{n_e} a_{i,j} }
\end{equation}

($a_i'$ becomes $a_i$ in the next computational step.)