### 2019-12-03 paper/briefpaper_stuedits/brief copy.tex
added sreamlined draft

% Brief report style paper

% It's about how a somatotopically ordered pattern can be
% re-established in the cortex based on information carried with
% thalamocortical axons afferent from the ventrobasal thalamus, but
% without specifying absolute positions as centre-points for the
% barrels. Everything in the paper should be tailored towards arguing
% this idea.

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
\templatetype{pnasresearcharticle}

%\title{Self-organization of whisker barrels in simulation requires competition between thalamocortical axons}
%\title{Simulated barrel cortex self-organizes using information carried by thalamocortical axons}
\title{Information carried by thalamocortical axons enables self-organization of the barrel field}
% While writing, don't stop for errors
%\nonstopmode

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Sebastian~S.~James}
\author[a]{Stuart~P.~Wilson}

\affil[a]{Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.}

% $^{*}$Corresponding author email: Seb.James@Sheffield.ac.uk\vspace{1em}\\

% Please give the surname of the lead author for the running footer
\leadauthor{James}

% Please add a significance statement to explain the relevance of your work
\significancestatement{The mystacial whiskers of rodents are a complex,
  spatially ordered set of structures whose pattern is repeated in several
  locations in the brain. We contribute a self-organising mechanism by which
  the pattern can be recreated with minimal genetic pre-determination at each
  location; only two pairs of axon guidance gradients are required in each
  tissue. The pattern's togography is carried by the axons which grow from one
  tissue to another and interact variably (depending upon which whisker they
  are associated with) with the guidance gradients in the target tissue whilst
  competing with each other for space. We demonstrate that this mechanism
  recreates a realistic whisker barrel field and also accounts for the results
  of well-known gene misexpression experiments.}

\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: \href{mailto:seb.james@sheffield.ac.uk}{\nolinkurl{seb.james@sheffield.ac.uk}}}

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
\keywords{Barrel cortex $|$ Self-organization $|$ Somatotopy $|$ Center-point $|$ Pattern}

\begin{abstract}
Understanding how genetic information and spontaneous pattern formation
together give rise to the whisker barrels is an important challenge for
developmental neurobiology. The barrel patterning that coalesces in layer IV
has been described formally as a Voronoi tesselation, which is indicative of
an underlying developmental process based on local competitive
interactions. If pattern formation relies only on such intrinsic
self-organising processes, how can the somatotopic order at the periphery be
maintained in the projections from brainstem to thalamus and from thalamus to
cortex?
%
To investigate, we began by extending a existing computational model of axon
branching and synaptogenesis in thalamocortical projections under the
influence of molecular signalling gradients.  We found that, although these
processes alone are incapable of generating patterns of cortical fields that
are more intricate than the patterns already present in the signalling
gradients, the addition of competition between the branches of axons that
originate from different thalamic barreloids allowed a full barrel-field
pattern to emerge in the presence of only two gradients.
%
%However, a
%full barrel-field pattern emerges under the guidance of as few as four
%signalling gradients when the model is extended to introduce competition
%between the branches of axons that originate from different thalamic
%barreloids.
%
%We show that the key requirements for the emergence of realistic
%(Voronoi tesselation) patterning are i) at each cortical location
%thalamocortical projections compete for a limited number of available
%connections, ii) at each location the axonal branching rate of each
%thalamocortical projection is reduced by the total branching density of the
%other projections, and iii) the density of axon branching for each
%thalamocortical projection is conserved.
%
We demonstrate how a purely
self-organising system can initialise a pattern which can then be faithfully
replicated in multiple neural structures.
%We show that this minimal model of
%barrel self-organisation provides a parsimonious account of data from a range
%of classic studies in which intrinsic and extrinsic factors have been
%experimentally manipulated.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Blue for comments
\newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
% Divergence symbol (Del dot Del)
\newcommand{\dvrg}{\nabla\vcdot\nabla}
% Emphasis shortcut
\newcommand{\e}{\emph}
% Bold shortcut
\newcommand{\bol}{\textbf}
% Math bold
\newcommand{\mb}[1]{\mathbf{#1}}
\makeatletter
\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
\makeatother

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\modulolinenumbers{}
\linenumbers

\dropcap{S}patial patterns in neural connectivity provide clues about the
constraints under which brains evolve and develop.
%Whether or not patterning
%governs information processing, models of the development of information
%processing circuits can be validated in terms of their ability to account for
%biological patterning \citep{purves_iterated_1992}.
%%see also \citealp{wilson_what_2015,bednar_cortical_2016}
One of the most
distinctive patterns of neural connectivity is found in the rodent
barrel cortex \cite{woolsey_structural_1970}.
%(\citealp{woolsey_pattern_1948,woolsey_structural_1970,welker_structure_1974},
%see \citealp{fox_barrel_2008}).
Understanding how this particular pattern forms might shed light on the
interplay of genetic and extrinsic factors that shape cortical circuits.
%, and on a wealth of
%anatomical, physiological, and behavioural data obtained using this
%model system.

The barrels are large cortical columns, identifiable from birth in layer 4 of
primary somatosensory cortex as dense clusters of thalamocortical axons, which
are enclosed by borders a few neurons thick from postnatal day 3
\citep{erzurumlu_development_2012}. In the plane tangential to the cortical
surface the barrels constitute a map of the arrangement of the whiskers on the
face,
%(see \citealp{yamakado_subdivision_1979})
such that the cells of adjacent barrels respond most strongly and most quickly
to deflection of adjacent whiskers \citep{armstrong-james_flow_1992}. The
barrels reflect subcortical whisker maps comprising cellular aggregates known
as barreloids in the thalamus \cite{van_der_loos_barreloids_1976} and
barrelettes in the brainstem.
%\citep{killackey_pattern_1980}.

%Each whisker and its corresponding barrelette, barreloid, and barrel can be
%identified by its position on a two-dimensional grid.
%\citep{zucker_coding_1969,killackey_pattern_1980,van_der_loos_barreloids_1976,haidarliu_size_2001}.
%
%The barrels are normally up to twice as long in the axis aligned to the grid
%‘rows’ compared to the ‘arcs’, and the rows are visible from day 2
%\citep{rebsam_refinement_2002} whereas the arcs become distinct by day 3
%\citep{erzurumlu_development_2012,rebsam_refinement_2002}.
%In metabotropic glutamate receptor 5 (mGlu5) knockouts the barrel rows develop
%but the arcs do not (\citealp{hannan_plc-1_2001} see also
%\citealp{fox_barrel_2008}), thus distinct processes likely guide development
%of the barrels with respect to the orthogonal grid axes.

Barrel formation requires afferent input, driven by whisker stimulation and by
thalamic calcium waves that propagate via mGlu5-receptor activation
\citep{anton-bolanos_prenatal_2019}.
%(\citealp{anton-bolanos_prenatal_2019}, see also
%\citealp{anton-bolanos_developmental_2018}).
Patterning also depends on a complex network of axon guidance processes, which
includes interactions between guidance molecules like ephrin-A5 and A7
\citep{miller_epha7-ephrin-a5_2006} and between adhesion molecules like
cadherin-6 and 8 \citep{bishop_regulation_2000}, and the
binding of p75 to neurotropins
\citep{bishop_distinct_2002}.
% or both citations?
%\citep{shimogori_fibroblast_2005,bishop_distinct_2002}.
%
%(\citealp{shimogori_fibroblast_2005,bishop_distinct_2002}
%; see also \citealp{dye_lifespan_2011,dye_lifespan_2011-1}).
This network is orchestrated by interactions between morphogens Fgf8 and Fgf17
and transcription factors Emx2, Pax6, Sp8, and Coup-tf1
\citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed
in gradients that mark orthogonal axes and can be manipulated to cause the
barrels to stretch, shrink, shift, skew, and even duplicate \cite{assimacopoulos_fibroblast_2012}.
% A longer list of citations:
%\citep{shimogori_fibroblast_2005,assimacopoulos_fibroblast_2012,borello_sp8_2014,sahara_sp8_2007,ypsilanti_transcriptional_2016,sur_patterning_2005}.
%
While complex structure in protein expression can be observed for some genes,
such as the molecules Bhlhb5 \citep{joshi_bhlhb5_2008}, we are not aware of
any report of gene expression patterns with the complexity of the barrel field
and so the mechanism by which the barrels form seems likely to be one in which
the structure forms by self-organization.

The boundaries of the barrels form a Voronoi tesselation across the cortical
sheet \citep{senft_mouse_1991}. This patterning suggests that the barreloid
topology may be preserved in the initial projection of the thalamocortical
axons into the cortex, and that a barrel forms by the lateral branching of the
axons from their initial centres of mass, which ceases upon contact with the
axons branching from adjacent centres. According to this model the barrel
pattern could be entirely determined by the distribution of the initial centre
points, which would itself be determined by guidance cues. However, the
assumption of pre-arranged centres is difficult to resolve with the
observation that thalamocortical axons arrive in the cortical plate as an
undifferentiated bundle prior to barreloid formation within the thalamus
\cite{agmon_organized_1993}. % Plus erzurumli, probably

Alternatively, reaction-diffusion dynamics could generate a Voronoi-like
tesselation in the cortex without the need for pre-arranged centres, by
amplifying characteristic modes in the noisy initial distribution of bundled
axon branches, as a net effect of short-range cooperative and longer-range
competitive interactions between cells. According to this model the barrel
pattern would be determined by the relative strength of these interactions and
by the shape of the cortical field boundary. However, reaction-diffusion
dynamics alone are unlikely to be able to account for the precise
co-registration and topographic connectivity between emergent domains in
thalamus and cortex, the non-uniform sizes and spacing of the barrels to form
rows versus arcs, and the influence of gene expression gradients.

As theories of barrel patterning, the centre-point and reaction-diffusion
models are not mutually exclusive. Pre-organised centres in the initial
conditions could bias reaction-diffusion processes to generate the specific
somatotopic arrangement of barrels more reliably, and similarly the mechanisms
by which axons branch laterally from barrel centres may involve a tension
between cooperative and competitive interactions. However, a proof that barrel
patterning can in principle emerge within an undifferentiated bundle of
thalamocortical axons, based only on local interactions between cells, would
remove the need to invoke a global supervisor, i.e., to assume a separate
stage and/or extrinsic mechanism for pre-organising thalamocortical
connections.

We seek to establish whether it is possible for the observed arrangement of
barrels to self-organise in a system governed by reaction-diffusion dynamics,
under the guidance of molecular signalling gradients, and in the absence of
pre-defined centre-points.
%
Our approach is to represent in a computational model the minimal set of
biological assumptions that might lead to the emergence of barrel patterning
in simulation, and, if necessary, to add to this minimal set only assumptions
that can be represented by the local exchange of information between adjacent
cells. The emergence of barrel patterning under such conditions would serve as
an existence proof of the plausibility of the hypothesis that barrel
development is a self-organising process.

\section*{Models}

Karbowski \& Ermentrout \citep{karbowski_model_2004} developed a dynamical
model of cortical arealization capable of generating predictions about the
influence of signalling molecules on the size, shape, and topological
organization of emergent cortical fields. Their model describes competition
between thalamocortical projections for territory along an anterior-posterior
cortical axis. The equations of the model define how the density of
connections, $c(x,t)$, and the density of axon branches, $a(x,t)$, interact at time $t$
at each cortical location $x$, for $N$ thalamocortical projections indexed by
$i$.
%The Voronoi tesselation used to describe barrel patterning is a
%two-dimensional structure, hence we first extended the original model for
%simulation on a two-dimensional cortical sheet, and describe it here as such
%(i.e., substituting $x$ with the two-dimensional position vector $\mb{x}$ and
%substituting terms of the form $\frac{d}{dx}$ with the spatial gradient
%operator $\nabla$).
%
The model was derived from the assumption that the rates at which
thalamocortical synapses form and thalamocortical axons branch are
reciprocally coupled, resulting in expressions for the change in $c_i$ as a
function of $a_i$, and for the change in $a_i$ as a function of
$c_i$. Synaptogenesis is modelled [in two dimensions, so the variables are
$a_i(\mb{x},t)$ \& $c_i(\mb{x},t)$] as
%
\begin{equation} \label{eq:dc}
\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{i'=1}^{N} c_{i'}\right)[a_i]^k.
\end{equation}
%
When the final term evaluates to zero, the density of thalamocortical
connections decays exponentially at a rate given by $\alpha$. This occurs when
the density of connections over all thalamic projections sums to one;
interpreted as a state in which all available cortical synaptic receptors are
occupied. Otherwise, the connection density is assumed to vary non-linearly
($k>1$) with the axon branching density.

Changes in axon branching density are assumed to depend also on the influence
of guidance molecules and on the density of branches at adjacent cortical
locations:
%
\begin{equation} \label{eq:da}
\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i}{\partial t} + \chi.
\end{equation}
%
The first term on the right describes the divergence (indicated by
$\nabla\vcdot$) of a quantity referred to as the `flux of axonal branching'
(the term enclosed by parentheses, called $\mb{J}_i(\mb{x},t)$).
%The central idea in heat diffusion
%models is that the change in temperature (over time) at a given location is
%proportional to the gradient in the temperature (over space) at that
%location. Substituting the branching density for heat, axons from a given
The flux consists of lateral diffusion across the cortical sheet at a rate
given by the diffusion constant $D$, and an interaction term, in which
branching is influenced by gradients in the concentration of $M$ distinct
molecular signals, $\rho(\mb{x})$. The influence of a given signal (indexed by
$j$) on a given thalamic projection (indexed by $i$), is determined by the
parameter $\gamma_{i,j}$ which may be positive or negative such that axons of
a given thalamic projection may branch preferentially towards either higher or
lower concentrations of each signalling molecule.
%Interactions for
%which this value is positive encourage axon branching in the direction in
%which the concentration of the signal is increasing, whereas negative
%interactions encourage branching in the opposite direction. Hence the axons of
%a given thalamic projection are assumed to branch towards either higher or
%lower concentrations of each signalling molecule.

The second term on the right of Eq.~\ref{eq:da} represents an assumption that the
axon branching density decreases when the number of connections being made
increases, and vice versa, i.e., that branches are required to form synaptic
connections. Note that we introduce the term $\chi=0$ here, so that we might
subsequently redefine $\chi$ to introduce competition between thalamic
projections.
%that we will show to be necessary for the emergence of a
%somatotopic pattern.

Two points are important to emphasise: First, notice that the only
difference between thalamic projections are the interaction strengths,
$\gamma_{i,j}$. Any differences between the size, shape, and location of
cortical fields that reliably emerge from the dynamics of the model must be
due to differences in the choice of these values only; Second, the formulation
of the model is at this point entirely local, in the sense that simulating the
dynamics requires no information to be communicated from cortical locations to
any but the immediately adjacent locations.

In addition to the choice of initial conditions (for the fields
$c(\mb{x},t=0)$ and $a(\mb{x},t=0)$) and the axon guidance fields
$\rho(\mb{x})$, the free parameters of the model are the number of
thalamocortical projections, $N$, constants $\alpha$, $\beta$, $k$, and $D$,
and the $N\times M$ interaction strengths, $\gamma$.

The dynamics of the model (Eqs.~\ref{eq:dc} \& \ref{eq:da}) were solved on a
hexagonal lattice of grid elements representing the cortical sheet, under the
assumption that axon branches cannot extend beyond the boundary of the sheet,
i.e., $\mb{J}_i(\mb{x},t) \big\rvert_{boundary} = 0$. We developed software
tools for imposing arbitrary boundary shapes on the lattice. The dynamics were
numerically integrated using fourth-order Runge-Kutta, and using a
{\color{red} finite difference method} to compute the flux term (see
Supplementary Materials).

\section*{Results}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{./MainFig.png}
\end{center}
%\vspace{-0em}
\caption{\textbf{A.} Barreloid centres taken from \cite{haidarliu_size_2001}
  are shown as black circles. Barreloids are labelled according to the usual
  convention. Labels R1 to R5 are the rhinal whisker barreloids. The dashed
  grey line represents the boundary of the barreloid field; the dot-dash line
  shows the anterior-posterior axis. It is assumed that there are two linearly
  increasing, orthogonal signalling gradients (molecules T1 \& T2) in the
  thalamus (side graphs), such that the location of each barreloid within the
  graded thalamic signal field can differentiate the efferent axons, with the
  result that axons from each barreloid interact to varying degrees with
  cortical signalling gradients (shown in panel B). Coloured arrows show, for
  each barreloid, the magnitude (and sign) of the interaction that its
  thalamocortical projection will have with the signalling gradients in the
  cortex. The mappings from barreloid location to interaction strength are
  linear. \textbf{B.} Guidance molecules in the simulated cortex. Molecule 1
  is expressed high to low in an approximately anterior to posterior
  gradient. Molecule 2 is expressed lateral to medial (analogously to ephrin -
  Vanderhaegen et al) \textbf{C.}  Whisker barrels from
  \cite{shimogori_fibroblast_2005}, with annotations for barrel identity and
  approximate axes.  \textbf{D-F.} Barrel cortex simulation. The maps indicate
  which thalomocortical axons make maximum connection density. Each hex
  element is coloured one of 52 unique colours depending on which e
  thalamocortical axon makes the most connections in that element. The mapping
  between barrel identity and colour is given in F.  \textbf{G.} The Honda
  Dirichlet measure as a function of simluation time for the simulations show
  in D-F. \textbf{H.}  Guidance molecules in the simulated Fgf8 misexpression
  experiment. Molecule 1 is mirrored over the anterior-posterior axis
  \textbf{I.} The simulation indicates that the mirroring of the barrel field
  can occur as a result of the mirroring of guidance molecule 1.}
\label{fig:main}
\end{figure*}

%\cmnt{If we need to present the results of the 2-D system with $\chi=0$, then
%  we'll do that first, showing a 2-D version of the simulation obtained by
%  running \code{james0 configs/karb.json}}

We first verified that the system given by Eqs.~\ref{eq:dc} \& \ref{eq:da}
described a two dimensional extension of the axon branching model in
\cite{karbowski_model_2004}. We showed that in the presence of three guidance
gradients varying only in one dimension (along the long axis of an elliptical
domain), five thalamocortical axon types, each interacting uniquely with the
three guidance gradients, would segregate into five separate regions. We
investigated a number of configurations of the guidance gradients, varying
their shape, $\rho(\mb{x})$, and number, $M$, and their effect on a varying
number of thalamocortical projections, $N$. In general, we were not able to
produce a complexity of structure in the resulting patterns of the
connections, $c(\mb{x},t\rightarrow\infty)$, which was not present in the
guidance patterns.

\subsection*{Branching competition is required to form somatotopically ordered barrels}

To investigate how the whisker barrel structure (Fig.~\ref{fig:main}.\,C) might
form in the presence of only monotonically varying morphogen gradients, we
turned our attention to the competition for connections built into the
model. There \emph{does} exist an element of competition in the original model
(Eqs.~\ref{eq:dc} \& \ref{eq:da}). The term $\left(1 - \sum_{i=1}^{N}
c_i\right)$ in Eq.~\ref{eq:dc} ensures that once the available synaptic
receptors in a region have been fully occupied by one thalamocortical axon
type, there is no longer the possibility for other types to form
synapses. This competition turns out to be relatively weak, because it does
not strongly affect the branching density in that region; it is possible for
two or more TC axon types to have roughly equal branching density in some
region, leading to roughly equal numbers of connections forming for each TC
type, up until the `hard' connection limit is reached.

To introduce competition between the \emph{branching} of thalamocortical axon
types, we introduced the new term $\chi$ given by
%
\begin{equation} \label{eq:comp}
\chi(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{i' \ne i}^{N} a_{i'}
\end{equation}
%
%with the choice $f(a) = a$,
which leads to a reduction in the rate of increase
of branching density for axon type $i$ in regions where the branching density
of \emph{other} axon types is high. The new parameter $\epsilon$ governs the
strength of the branching competition and the factor $1/(N-1)$ ensures that the
strength scales with $N$.

With the addition of branching competition, it is necessary to introduce a
normalisation step, in order to prevent one TC type from out-competing every
other type. This represents the idea that for a fixed number of afferent
thalamocortical axons, the overall number of branches must have some upper
limit, given the finite resources available in each axon for building new
processes. We divisively normalize the branching density, dividing it at each
computational step by the mean branching density per unit area. It is then
multiplied by the branching density per unit area at time 0, to account for
the possibility that there might be more axons afferent for TC type $1$ cf.~TC
type $2$:
%
\begin{equation} \label{eq:norm}
  a_i(\mathbf{x}, t) = \int_\Omega  a_i(\mathbf{x}, 0) d\sigma \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
  a'_i(\mathbf{x}, t) d\sigma}
\end{equation}
%
The prime symbols indicate the use of the intermediate value of $a_i$ after it
has been computed according to Eqs.~\ref{eq:dc}--\ref{eq:comp} on the
basis of the system state at time $t-{\delta}t$.

\subsection*{The model can transfer realistic barrel patterns}

With these changes we were able to generate grid-like patterns with
appropriate (linearly spaced) choices of the interaction parameters,
$\gamma$. To determine the $\gamma$ values suitable to recreate the cortical
barrel field by a principled method, we postulated that the interaction of
each thalamocortical axon group with the ephrin gradients in the cortical
sub-plate might be determined by the location of the barreloid within
signalling gradents \emph{in the thalamus}. We therefore sought a map of the
thalamic barreloids from the literature. \cite{haidarliu_size_2001} provided a
comprehensive map of the layout of the barreloids in rat, which is similar to
mouse \cite{van_der_loos_barreloids_1976}. We used these data to create the
map of 52 barreloids shown in Fig.~\ref{fig:main}A. We postulated two linearly
varying molecules and used the barreloids' positions within this field to
generate the interaction parameters represented by the blue and red arrows.

We configured a simulation based on Eqs.~\ref{eq:dc}--\ref{eq:norm} to run on
a domain whose shape was matched to data from \cite{shimogori_fibroblast_2005}
(Fig.~\ref{fig:main}A). 52 separate thalamocortical fields were initialised
with random noise in the branching density and given our thalamic interaction
parameters. Two orthogonal gradients were specified; molecule 1 with an
posterior to anterior gradient and molecule 2 resembling ephrin A5, with a
medial-lateral gradient. Fig.~\ref{fig:main}D--E shows snapshots of the
connection density at simulation times x (800 steps), y (2000) and z (3200).

% On stronger branching; not used in the results presented here.
% \cmnt{The realistic barrel pattern experiment is enhanced by stronger
%  branching competition:}
%
% To increase the strength of the competition, we choose an alternative form for
% $f(a)$ in Eq.~\ref{eq:comp} which has a geometric form for small $a$: $f(a) =
% \frac{1}{1 + e^{-m[a]^l}} - \frac{1}{2}$. The parameters $l$ and $m$ control
% the steepness and shape of the function.
%%% End of stronger branching competition %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{The barrel field reacts to changes in the expression pattern of guidance molecules}

To cover the reaction to changes in the guidance molecules caused by
experiments like those by Shimogori \& Grove.

The code and configuration required to reproduce the results in this paper are
available at \url{https://github.com/ABRG-Models/BarrelEmerge}.

\section*{Discussion}

We set out to test the hypothesis that barrel patterning may emerge via
self-organising processes. We first extended an existing model of neocortical
arealization to simulate pattern formation in two dimensions, inheriting from
this model the assumption that axonal branching and synaptogenesis are
coupled, and that thalamic projections compete for cortical territory under
the influence of molecular signalling gradients. Realistic barrel patterning
emerged only when terms were added to represent the assumption that
thalamocortical projections compete for a limited supply of connections with
cortical cells, that their branching rates are reduced by the branching of
other thalamocortical projections, and that the density of axon branching is
conserved for each projection. The net effect of these local interactions is a
Voronoi tesselation in which localised clusters of thalamic projections become
somatotopically organised on the cortical sheet according to the relative
values of coefficients specifying the rates at which axons are attracted and
repelled by orthogonal pairs of molecular signal gradients. According to this
model, the pattern of barrels is determined by the competitive and cooperative
interactions between cells relative to the shape of the cortical field
boundary. In simulation, we systematically investigated the origin of the
asymmetry between the barrel rows and arcs, and identified {\color{red} XXXXX}
as the most likely candidate for the elongation of barrel shapes along the
rows.

\bibliography{../BarrelEmerge}

\end{document}