### 2019-12-03 paper/briefpaper_stuedits/brief.tex

% Brief report style paper

% It's about how a somatotopically ordered pattern can be
% re-established in the cortex based on information carried with
% thalamocortical axons afferent from the ventrobasal thalamus, but
% without specifying absolute positions as centre-points for the
% barrels. Everything in the paper should be tailored towards arguing
% this idea.

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
\templatetype{pnasresearcharticle}

\title{Modelling the emergence of whisker barrels}
% While writing, don't stop for errors
%\nonstopmode

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Sebastian~S.~James}
\author[a]{Stuart~P.~Wilson}

\affil[a]{Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.}

% $^{*}$Corresponding author email: Seb.James@Sheffield.ac.uk\vspace{1em}\\

\leadauthor{James}

\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: \href{mailto:seb.james@sheffield.ac.uk}{\nolinkurl{seb.james@sheffield.ac.uk}}}

\keywords{Barrel cortex $|$ Self-organization $|$ Somatotopic map $|$ Axon guidance}

\begin{abstract}
Brain development relies on an interplay of genetic specification and self-organisation. Striking examples of this interplay can be found in the rodent somatosensory brainstem, thalamus, and cortex, where the arrangement of the facial whiskers is preserved in the arrangement of cellular aggregates to form precise somatotopic maps. We show in simulation how realistic whisker maps can self-organize, by assuming that information is exchanged between adjacent cells only, under the guidance of gene expression gradients. The resulting model provides a simple account of how genetic information can constrain spontaneous pattern formation to faithfully reproduce functional maps in subsequent brain structures.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Blue for comments
\newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
\newcommand{\dvrg}{\nabla\vcdot\nabla}
\newcommand{\e}{\emph}
\newcommand{\bol}{\textbf}
\newcommand{\mb}[1]{\mathbf{#1}}
\makeatletter
\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
\makeatother

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\modulolinenumbers{}
\linenumbers

\dropcap{S}patial patterns in neural connectivity provide clues about the constraints under which brains evolve and develop \citep{purves_iterated_1992}. Perhaps the most distinctive pattern of neural connectivity can be found in the rodent barrel cortex \cite{woolsey_structural_1970}. The barrels are large cortical columns, identifiable from birth in layer 4 of primary somatosensory cortex as dense clusters of thalamocortical axons, which are enclosed by borders a few neurons thick from postnatal day 3 \citep{erzurumlu_development_2012}. In the plane tangential to the cortical surface the barrels constitute a somatotopic map of the whiskers, with cells of adjacent barrels responding most strongly and quickly to deflection of adjacent whiskers \citep{armstrong-james_flow_1992}. Barrel patterning reflects subcortical whisker maps comprising cell aggregates called barrelettes in the brainstem and barreloids in the thalamus \cite{van_der_loos_barreloids_1976}.

Barrel patterning requires afferent input from whisker stimulation and thalamic calcium waves \citep{anton-bolanos_prenatal_2019}, and depends on a complex network of axon guidance molecules such as ephrin-A5 and A7 and adhesion molecules such as cadherin-6 and 8 \citep{miller_epha7-ephrin-a5_2006}.
This network is orchestrated by interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2, Pax6, Sp8, and Coup-tf1 \citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed in gradients that mark orthogonal axes and can be manipulated to stretch, shrink, shift, and even duplicate barrels \cite{assimacopoulos_fibroblast_2012}.

The barrel boundaries form a Voronoi tessellation across the cortical sheet \citep{senft_mouse_1991}, suggesting that barreloid topology is preserved in the projection of thalamocortical axons into the cortex, and that a barrel forms by lateral axon branching from an initial centre-point that ceases upon contact with axons branching from adjacent centres. But the assumption of pre-arranged centres is difficult to resolve with the observation that  axons arrive in the cortical plate as an undifferentiated bundle, prior to barreloid formation \cite{agmon_organized_1993}.

Alternatively, reaction-diffusion dynamics could generate a Voronoi tessellation without pre-arranged centres, by amplifying characteristic modes in a noisy initial distribution of axon branches, as a net effect of short-range cooperative and longer-range competitive interactions. Accordingly, the barrel pattern would be determined by the relative strength of these interactions and by the shape of the cortical field boundary. However, intrinsic cortical dynamics alone cannot account for the topographic correspondence between domains in the thalamus and cortex, or the irregular sizes and specific arrangement of the barrels in rows and arcs, or the influence of gene expression gradients.

The centre-point and reaction-diffusion models are not mutually exclusive. Pre-organised centres could bias reaction-diffusion processes to generate specific  arrangements more reliably, and mechanisms of lateral axon branching may constitute the tension between cooperation and competition required for self-organization. However, proof that barrel patterning can emerge from an undifferentiated bundle of  axons, based only on local interactions, would show that a separate stage and/or extrinsic mechanism for pre-organising thalamocortical connections need not be assumed. To this end, we ask whether barrel maps can emerge in a system with reaction-diffusion dynamics, under the guidance of signalling gradients, and in the absence of pre-defined centres.

\section*{Models}

Karbowski \& Ermentrout \citep{karbowski_model_2004} developed a reaction-diffusion style model of how extrinsic signalling gradients can constrain the emergence of distinct fields from intrinsic cortical dynamics. Their model defines how the density of connections $c(x,t)$, and axon branches $a(x,t)$, interact at time $t$, along a 1D anterior-posterior axis $x$, for $N$ thalamocortical projections indexed by $i$. The model was derived from the assumption that the rates at which $a_i$ and $c_i$ grow are reciprocally coupled. Extending the original 1D model to simulate arealization on a 2D cortical sheet, we use $a_i(\mb{x},t)$ and $c_i(\mb{x},t)$, and  model synaptogenesis as
%
\begin{equation} \label{eq:dc}
\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{k=1}^{N} c_{k}\right)[a_i]^l.
\end{equation}
%
When the final term evaluates to zero, i.e., when the total density of connections sums to one, the density of thalamocortical connections decays at rate $\alpha$. Otherwise the synaptic connection density increases non-linearly ($l>1$) with the axon branching density. Changes in axon branching are modelled as 
%
\begin{equation} \label{eq:da}
\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i}{\partial t} + \chi_i.
\end{equation}
%
The first term on the right describes the divergence (indicated by $\nabla\vcdot$) of the quantity in parentheses, which is referred to as the `flux of axonal branching'. The flux represents diffusion across the cortical sheet, at rate $D$, and the influence of $M$ molecular signalling fields, $\rho(\mb{x})$. The influence of a given field (indexed by $j$) on a given thalamic projection (indexed by $i$), is determined by $\gamma_{i,j}$, which may be positive or negative in order that axons may branch in the direction of either higher or lower  concentrations. The second term on the right represents an assumption that axon branching decreases when synaptogenesis increases, and vice versa, i.e., that axon branches are required for synaptic connections to form. Here $\chi_i=0$ is a placeholder.

Software tools for modelling the cortical sheet as a hexagonal lattice with no-flux boundary conditions and an arbitrary boundary shape, for computing the flux using a finite difference method, for numerical integration by the fourth-order Runge-Kutta method, and for reproducing the reported results are available at \url{https://github.com/ABRG-Models/BarrelEmerge}.

\section*{Results}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{./MainFig.png}
\end{center}

\caption{\textbf{A.} Barreloid centres taken from \cite{haidarliu_size_2001} are shown as black circles. Barreloids are labelled according to the usual convention. Labels R1 to R5 are the rhinal whisker barreloids. The dashed grey line represents the boundary of the barreloid field; the dot-dash line shows the anterior-posterior axis. It is assumed that there are two linearly increasing, orthogonal signalling gradients (molecules T1 \& T2) in the thalamus (side graphs), such that the location of each barreloid within the graded thalamic signal field can differentiate the efferent axons, with the result that axons from each barreloid interact to varying degrees with cortical signalling gradients (shown in panel B). Coloured arrows show, for each barreloid, the magnitude (and sign) of the interaction that its thalamocortical projection will have with the signalling gradients in the cortex. The mappings from barreloid location to interaction strength are linear. \textbf{B.} Guidance molecules in the simulated cortex. Molecule 1 is expressed high to low in an approximately anterior to posterior gradient. Molecule 2 is expressed lateral to medial (analogously to ephrin - Vanderhaegen et al) \textbf{C.}  Whisker barrels from \cite{shimogori_fibroblast_2005}, with annotations for barrel identity and approximate axes.  \textbf{D-F.} Barrel cortex simulation. The maps indicate which thalomocortical axons make maximum connection density. Each hex element is coloured one of 52 unique colours depending on which thalamocortical axon makes the most connections in that element. The mapping between barrel identity and colour is given in F.  \textbf{G.} The Honda Dirichlet measure as a function of simluation time for the simulations show in D-F. \textbf{H.}  Guidance molecules in the simulated Fgf8 misexpression experiment. Molecule 1 is mirrored over the anterior-posterior axis \textbf{I.} The simulation indicates that the mirroring of the barrel field can occur as a result of the mirroring of guidance molecule 1.}
\label{fig:main}
\end{figure*}

First we verified that all results established for a 1D axis could be reproduced using our extension to a 2D cortical sheet. Using an elliptical boundary with $M=3$ guidance gradients aligned to the longer axis, $N=5$ thalamocortical projections gave rise to five distinct cortical fields at locations that preserved the topographic ordering defined by the original $\gamma$ values of \cite{karbowski_model_2004}. However, we found that specifying $N$ ordered areas required $M\approx N$ signalling fields. As the number of distinct guidance fields is unlikely to approach the number of barrels, further extensions were required.

The term in parentheses in Eq.~\ref{eq:dc} represents competition between thalamic projections for a limited availability of cortical connections. To introduce competition also in terms of axon branching we introduced
%
\begin{equation} \label{eq:comp}
\chi_i(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{j \ne i}^{N} a_{j},
\end{equation}
%
which reduces branching for each projection where the branching of other projections is high. Divisive normalization is used to keep the axon branch density bounded on each iteration to the initial density  at $t=0$,

\begin{equation} \label{eq:norm}
  a_i(\mathbf{x}, t) = \int_\Omega  a_i(\mathbf{x}, 0) d\sigma \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
  a'_i(\mathbf{x}, t) d\sigma},
\end{equation}
%
where prime symbols indicate the use of intermediate values computed from Eqs.~\ref{eq:dc}--\ref{eq:comp} at time $t-{\delta}t$.

The only differences between thalamic projections are the interaction strengths, $\gamma$, hence any reliable differences between emergent cortical fields must be due to differences in these values only. We chose $\gamma$ values by speculating that interaction strengths with ephrin gradients in the cortical subplate are determined by sampling from similar thalamic gradients at the origin of each thalamic projection, i.e., at each barreloid centre. Two orthogonal linear thalamic gradients were defined, from which 52 pairs of $\gamma$ values were sampled, at the coordinates of 52 barreloid centers recreated from  \cite{haidarliu_size_2001}.

A cortical boundary enclosing 52 corresponding barrels was traced from a cytochrome oxidase stain from  \cite{shimogori_fibroblast_2005}, and Eqs.~\ref{eq:dc}--\ref{eq:norm} were solved for $N=52$ projections on the resulting domain, using $M=2$ linear signalling gradients aligned with the anterior-posterior and  medial-lateral axes. From random initial conditions for $a({\bf x},0)$ and $c({\bf x},0)$, a clear Voronoi-like tesselation emerged. A reduction in the Honda delta metric (see \cite{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged within $\approx 3000$ iterations (see Figure 1). The sum of squared differences between the real and simulated barrel centres revealed a strong correspondence.

To further investigate the interplay of genetic and intrinsic mechanisms we simulated a seminal barrel duplication experiment \cite{shimogori_fibroblast_2005},  first by solving on a cortical domain comprising two separate, identically shaped cortical boundaries. In this case one distinct cortical field per thalamic projection emerged, within one of the two boundaries only, i.e., no duplication. However, merging the two fields to create an extended mirror-symmetric boundary shape, with an anterior-posterior guidance field reflected at the join, gave rise to two mirror-symmetrical barrel fields comprising 104 barrels, i.e., two cortical fields for each thalamic projection (see Figure 2). Together these results suggest that by the misexpression of Fgf8, Shimogori \& Grove \citep{shimogori_fibroblast_2005} effectively created one large barrel field rather than two distinct fields.

\section*{Discussion}

The present results suggest that the key requirements for the emergence of realistic barrel patterning are i) at each cortical location thalamocortical projections compete for a limited number of available synaptic connections (Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a given projection is reduced by the density of other projections (Eqs.~\ref{eq:comp}), and iii) the branch density of each projection is conserved (Eq.~\ref{eq:norm}).

It is important to emphasize that the formulation of the model is entirely local, insofar as simulation requires no information to be communicated between thalamic projections, or from a given cortical location to any but those immediately adjacent. Hence the simulations demonstrate how a self-organising system, constrained by genetically specified guidance cues, can faithfully reproduce an arrangement of cell aggregates in one neural structure as a topographic map in another.

The model also shows that the shape of the barrels may be primarily determined by the competitive and cooperative interactions between cells relative to the shape of the cortical field boundary. And that somatotopic map formation does not require the pre-specification of centre-points by as yet undetermined additional developmental mechanisms.

\bibliography{../BarrelEmerge}

\end{document}