### 2020-06-30 paper/briefpaper/brief.tex
Colour wrangling, label changing

% Brief report style paper

% It's about how a somatotopically ordered pattern can be
% re-established in the cortex based on information carried with
% thalamocortical axons afferent from the ventrobasal thalamus, but
% without specifying absolute positions as center-points for the
% barrels. Everything in the paper should be tailored towards arguing
% this idea.

\documentclass[9pt,lineno]{elife}

\usepackage{color}
\usepackage{amsmath,esint}

% Some document-defined commands
\newcommand*\dif{\mathop{}\!\mathrm{d}}

% Colours for revisions:
% Custom colours
\definecolor{revgreen}{rgb}{0.220, 0.463, 0.114}
\definecolor{revblue}{rgb}{0.235, 0.471, 0.847}
\definecolor{revred}{rgb}{0.8353, 0.1255, 0.1255}
\definecolor{revorange}{rgb}{0.8235, 0.4275, 0.0}
\definecolor{revgrey}{rgb}{0.6, 0.6, 0.63}
\definecolor{revcomp}{rgb}{0.0, 0.0, 0.0}
\definecolor{revpurp}{rgb}{0.8, 0.1, 0.8}
\definecolor{revpurp2}{rgb}{0.3882, 0.0745, 0.7098}
\definecolor{revpink}{rgb}{0.7098, 0.0745, 0.6431}
\definecolor{revbblack}{rgb}{0, 0, 0.3}
\definecolor{revgblack}{rgb}{0, 0.2, 0}

% Miscellaneous changes
\newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
% Justify gamma/interaction parameter choice. major point 1.
\newcommand{\MPone}[1]{\textcolor{revpink}{#1}}
% Blue changes are major point 2 (competition mechanism) and other, general
% uncategorised changes.
\newcommand{\MPtwo}[1]{\textcolor{revblue}{#1}}
% Parameter search. Major point 3
\newcommand{\MPthreePar}[1]{\textcolor{revbblack}{#1}}
% Sensitivity analysis. Major point 3
\newcommand{\MPthreeSens}[1]{\textcolor{revgblack}{#1}}
% Justify competition:
\newcommand{\mpfive}[1]{\textcolor{revpurp2}{#1}}
% Minor point 4:
\newcommand{\mpfour}[1]{\textcolor{revgreen}{#1}}
% Minor point 6:
\newcommand{\mpsix}[1]{\textcolor{revpurp}{#1}}
% Describe the metrics
\newcommand{\metrics}[1]{\textcolor{revred}{#1}}

\newcommand{\dvrg}{\nabla\vcdot\nabla}
\newcommand{\e}{\emph}
\newcommand{\bol}{\textbf}
\newcommand{\mb}[1]{\mathbf{#1}}
\makeatletter
\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
\newcommand{\code}[1]{\textsf{#1}}
% Nice norms:
\usepackage{mathtools}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
% Degrees symbol:
\usepackage{textcomp}

\title{Modelling the emergence of whisker barrels}

\author[1*]{Sebastian~S.~James}
\author[2]{Leah~A.~Krubitzer}
\author[1]{Stuart~P.~Wilson}

\affil[1]{Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.}
\affil[2]{Center for Neuroscience, The University of California, Davis, United States.}
\corr{seb.james@sheffield.ac.uk}{SSJ}

%\keywords{Barrel cortex $|$ Self-organization $|$ Somatotopic map $|$ Axon guidance}

\begin{document}

\maketitle

\begin{abstract}
Brain development relies on an interplay between genetic specification and
self-organization. Striking examples of this relationship can be found in the
somatosensory brainstem, thalamus, and cortex of rats and mice, where the
arrangement of the facial whiskers is preserved in the arrangement of cell
aggregates to form precise somatotopic maps. We show in simulation how
realistic whisker maps can self-organize, by assuming that information is
exchanged between adjacent cells only, under the guidance of gene expression
gradients. The resulting model provides a simple account of how patterns of
gene expression can constrain spontaneous pattern formation to faithfully
reproduce functional maps in subsequent brain structures.
\end{abstract}

\section{Introduction}

Spatial patterns in neural connectivity provide clues about the constraints
under which brains evolve and develop \citep{purves_iterated_1992}. Perhaps
the most distinctive pattern can be found in the barrel cortex of many rodent
species \citep{woolsey_structural_1970}. The barrels are identifiable soon
after birth in layer 4 of primary somatosensory cortex as dense clusters of
thalamocortical axons, which are enclosed by borders a few neurons thick from
postnatal day 3 \citep{erzurumlu_development_2012}.
%
% These timings, from erzurumlu_development_2012 originate from
% sehara_neuronal_2011, where they relate to mouse.
\mpsix{In the mouse, axons from the trigeminal ganglion arrive in the principal sensory
  nucleus of the trigeminal nerve (PrV) at E12, then axons from the PrV arrive
  in the ventroposteromedial nucleus of the thalamus (VPM) at E17, then axons
  from the VPM arrive in the cortical plate at E18/P0. Distinct whisker-related
  clusters then become apparent in the PrV at P0-P1, in the VPM at P2-P3, and
  in the cortex at P3-P5} \citep{erzurumlu_development_2012,sehara_neuronal_2011}.
%
In the plane tangential to
the cortical surface the barrels constitute a somatotopic map of the whiskers,
with cells within adjacent barrels responding most strongly and quickly to
deflection of adjacent whiskers \citep{armstrong-james_flow_1992}. Barrel
patterning reflects subcortical whisker maps comprising cell aggregates called
barrelettes in the brainstem and barreloids in the thalamus
\citep{ma_barrelettesarchitectonic_1991,van_der_loos_barreloids_1976}.

Barrel formation requires afferent input from whisker stimulation and thalamic
calcium waves \citep{anton-bolanos_prenatal_2019}, and depends on a complex
network of axon guidance molecules such as ephrin-A5 and A7 and adhesion
molecules such as cadherin-6 and 8
\citep{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}.  This network
is orchestrated by interactions between morphogens Fgf8 and Fgf17 and
transcription factors Emx2, Pax6, Sp8, and Coup-tf1
\citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed
in gradients \mpfour{across the full cortical sheet} that mark orthogonal axes
and can be manipulated to stretch,
shrink, shift, and even duplicate barrels
\citep{assimacopoulos_fibroblast_2012}.

The barrel boundaries form a Voronoi tessellation across the cortical sheet
\citep{senft_mouse_1991} (Fig.\,\ref{fig:main}A), suggesting that barreloid topology is
preserved in the projection of thalamocortical axons into the cortex, and that
a barrel forms by lateral axon branching from an initial center-point that
ceases upon contact with axons branching from adjacent centers.  However, the
assumption of pre-arranged center-points is difficult to resolve with the
observation that axons arrive in the cortical plate as an undifferentiated
bundle, \emph{prior} to barreloid formation \citep{agmon_organized_1993}.

% Note: The key difference between this theory and the chemospecificity theory
% is that of group membership.

Alternatively, reaction-diffusion dynamics could generate a Voronoi
tessellation without pre-arranged centers, by amplifying characteristic modes
in a noisy initial distribution of axon branches, as a net effect of
short-range cooperative and longer-range competitive
interactions. Accordingly, the barrel pattern would be determined by the
relative strength of these interactions and by the shape of the cortical field
boundary. However, intrinsic cortical dynamics alone cannot account for the
topographic correspondence between thalamic and cortical domains, the
irregular sizes and specific arrangement of the barrels in rows and arcs, or
the influence of gene expression gradients.

The center-point and reaction-diffusion models are not mutually
exclusive. Pre-organized centers could bias reaction-diffusion processes to
generate specific arrangements more reliably, and mechanisms of lateral axon
branching may constitute the tension between cooperation and competition
required for self-organization. However, proof that barrel patterning can
emerge from an undifferentiated bundle of axons, based only on local
interactions, would show that a separate stage and/or extrinsic mechanism for
pre-organizing thalamocortical connections need not be assumed. To this end,
we ask whether barrel maps can emerge in a system with reaction-diffusion
dynamics, under the guidance of signalling gradients, and in the absence of
pre-defined centers.

\section{Models}

\cite{karbowski_model_2004} developed a reaction-diffusion style model of how
extrinsic signalling gradients can constrain the emergence of distinct fields
from intrinsic cortical dynamics. Their model defines how the density of
connections $c(x,t)$ and axon branches $a(x,t)$ interact at time $t$, along a
1D anterior-posterior axis $x$, for $N$ thalamocortical projections indexed by
$i$. The model was derived from the assumption that the rates at which $a_i$
and $c_i$ grow are reciprocally coupled. Extending the original 1D model to
simulate arealization on a 2D cortical sheet, we use $a_i(\mb{x},t)$ and
$c_i(\mb{x},t)$, and model synaptogenesis as
%
\begin{equation} \label{eq:dc}
\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{j=1}^{N} c_{j}\right)[a_i]^k.
\end{equation}
%
Accordingly, where the total density of synaptic connections sums to one,
connections decay at rate $\alpha$. Otherwise the connection density increases
non-linearly ($k>1$) with the density of axon branching. Axon branching is
modelled as
%
\begin{equation} \label{eq:da}
\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \MPtwo{+ \chi_i}\right) - \frac{\partial c_i}{\partial t}.
\end{equation}
%
The first term on the right describes the divergence (indicated by
$\nabla\vcdot$) of the quantity in parentheses, which is referred to as the
`flux' of axonal branching. The flux represents diffusion across the cortical
sheet, at rate $D$, and the influence of $M$ molecular signalling fields,
$\rho(\mb{x})$. The influence of a given field (indexed by $j$) on a given
thalamic projection (indexed by $i$), is determined by $\gamma_{i,j}$, which
may be positive or negative in order that axons may branch in the direction of
either higher or lower concentrations. Note that computing the divergence in
simulation requires cells on the cortical sheet to communicate with immedately
adjacent cells only (see \emph{Materials \& Methods}). The second term on the
right quantifies the coupling between axon branching and synaptogenesis. Here
  $\chi_i=0$ is a placeholder.

\section{Results}

\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig1.png}
    \caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
      by \cite{zheng_signal_2001}, with black lines to delineate barrels and to
      measure departure (Honda-$\delta$; see \citealp{senft_mouse_1991}) from a
      perfect Voronoi tesselation. Right shows the initial distribution of axon
      branching density ($a$) for one thalamocortical projection, and two
      molecular guidance fields ($\rho$), \cmnt{where the domain $S$ has been
        traced from \textbf{A}}. \textbf{B} The strengths of interaction
      $\gamma$ with fields $\rho_1$ and $\rho_2$ are indicated for each of 41
      projections by the lengths of green and blue arrows respectively, assuming
      that similar fields aligned to the posterior-anterior and medial-lateral
      axes in the ventroposterior medial nucleus of the thalamus are sampled at
      the locations of putative barreloid centers (reconstructed from
      \citealp{haidarliu_size_2001}, \MPone{Fig.\,5b}). \textbf{C} Simulation results for parameters
      \MPtwo{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
      $\epsilon=1.2$} and $\delta{t}=0.0001$. Colours indicate the thalamic
      projection for which the connection density is maximal and black lines
      delineate boundaries (see Movie
      S1). \textbf{D} Red dots show the Honda metric, \metrics{$\delta(t)$}, obtained from \metrics{the}
      simulation approaching that obtained from \metrics{real} barrels in \textbf{A} (dotted
      line); black squares \metrics{provide a} measure \metrics{of} the correspondence
      between the real and
      simulated barrel shapes, $\eta(t)$ (the \metrics{pattern quality metric---see
        Materials \& Methods; units mm$^3$});
      \metrics{grey hexagons \metrics{provide a measure of how selectively each
          cortical site is innervated};
        $\omega(t) = \oiint_{S} \mu(\mb{x},t) \mathrm{d}S$, where
        $\mu(\mb{x}) \equiv \mathrm{max}_i\big(c_i(\mb{x},t)\big)\big/\sum_{j=1}^{N} c_j(\mb{x},t)$.}
      \textbf{E} \metrics{Plotted across the cortical sheet, the selectivity
        develops to reveal an alignment with the emergent barrel boundary
        shapes}. \metrics{The colour map
        indicates the values of $\mu(\mb{x})$. Overlaid
        contours show $c_i > 0.95 \mathrm{max}_i\big(c_i(\mb{x})\big)$.} All scale bars 1\,mm.}
    \label{fig:main}
  \end{fullwidth}
\end{figure}

First we verified that all results established by \cite{karbowski_model_2004}
for a 1D axis could be reproduced using our extension to a 2D cortical
sheet. Using an elliptical domain, $S$, with $M=3$ offset guidance gradients
aligned to the longer axis, $N=5$ thalamocortical projections gave rise to
five distinct cortical fields at locations that preserved the topographic
ordering defined by the original $\gamma$ values. However, we found that
specifying $N$ ordered areas required $M\approx (N+1)/2$ signalling
fields. This is because localization of axon densities occurs only when
projections are influenced by interactions with two or more signalling
gradients that encourage migration in opposing directions.
%
%Note that these dynamics are quite unlike classic chemospecificity models
%\citep{sperry_chemoaffinity_1963}, which essentially assume center-points
%i.e., conditions in the target tissue that instruct pre-identified afferents
%to stop growing.
%
As the number of distinct guidance fields is unlikely to
approach the number of barrels, \cmnt{modifications to the model} were required.

% Justify competition term. This pulled from Stuart's Google doc
\mpfive{We reasoned that an arbitrary number of distinct field locations
  may be specified using a minimum of two guidance fields, if the tendency for
  the density of each projection to concentrate is appropriately balanced
  against the tendency of projections that interact more strongly with the
  guidance fields to migrate further in the directions specified by their
  gradients. Accordingly, projections that interact most strongly with a given
  guidance gradient would come to occupy cortical locations at which that
  field has extreme values, leaving adjacent locations available to be
  occupied by projections with the next strongest interactions, and so
  forth. This would in principle allow the \emph{relative} locations of the
  fields to be specified by the relative values of the interaction parameters,
  $\gamma$, and hence for a topological map in the cortex to be specified by a
  spatial ordering of the $\gamma$ values at the level of the thalamus.}

\mpfive{Such dynamics are quite unlike those described by classic
  chemospecificity models} \citep{sperry_chemoaffinity_1963},
\mpfive{which essentially assume center-points by specifying conditions
  in the target tissue that instruct pre-identified afferents to stop
  growing. Consider, for example, that when simulated in isolation from
  one-another, all projections in the model described would simply migrate to
  the extrema of the cortical guidance fields.}

\mpfive{Testing this reasoning required increasing the strength of the
  competition between simulated thalamocortical projections for cortical
  territory, by increasing the tendency for the density of each projection to
  concentrate. The major modification required was thus to introduce into the
  model an additional source of competition between thalamic projections.}
%
The term in parentheses in Eq.\,\ref{eq:dc} represents competition between
thalamocortical projections for a limited availability of cortical
connections. To introduce competition also in terms of axon branching,
\MPtwo{whilst ensuring that $a_i$ is conserved over time,} we
redefined
%
\begin{equation} \label{eq:comp}
  \color{revblue}
  \chi_i(\mb{x}, t) = \frac{\epsilon a_i}{N-1} \nabla \textstyle\sum_{j\ne i}^{N}{a_j}.
  \color{black}
\end{equation}
%
\mpfive{This term contributes to the \emph{flux} of axonal branching as
  an additional source of diffusion, which reduces the branching density for a
  given projection where the branches of other projections are dense. The
  strength of this additional competition is parameterized by
  $\epsilon$}. Note that this operation is local to individual afferent
projections.

\MPone{In addition, the model outlined requires that molecular
  guidance gradients in the cortex are complemented by graded values of the
  interaction strengths, $\gamma$, at the level of the thalamus. While the
  precise mechanisms by which thalamic and cortical gradients interact during
  development have not been fully characterised, the presence of complementary
  thalamic and cortical molecular guidance gradients has been well established
  experimentally. In particular, the EphA4 receptor and its ligand ephrin-A5
  are distributed in complementary gradients in the somatosensory thalamus and
  cortex} (\citealp{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}).
\MPone{Cells originating in the medial part of ventrobasal thalamus
  (comprising VPM and VPL) express high levels of EphA receptors and project
  to the lateral part of S1, which expresses low levels of ephrin-A5, and
  cells originating in the lateral part of ventrobasal thalamus express low
  levels of EphA receptors and project to the medial part of S1, which
  expresses high levels of ephrin-A5} (see
\citealp{gao_regulation_1998,dufour_area_2003,vanderhaeghen_developmental_2004,speer_grading_2005,torii_role_2013}).

\MPone{For simplicity, both the thalamic interaction gradients and the
  cortical guidance fields in the model were chosen in the first instance to
  be linear and orthogonal. Hence a given pair of gamma values corresponds to
  the coordinate of a barreloid center in the VPM (in a reference plane
  defined by the anterior-posterior and medial-lateral axes. Coordinates were
  estimated from Fig 5d of} \citep{haidarliu_size_2001}, \MPone{and
  scaled (arbitrarily) such that $\gamma\in\pm 2$.}

A cortical boundary enclosing \cmnt{the} 41 \cmnt{macrovibrissae} barrels was
traced from a cytochrome oxidase stain from \cite{zheng_signal_2001}
\cmnt{(using original data kindly supplied by the authors)}, and
Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved for $N=41$ projections on the
resulting domain, using $M=2$ linear signalling gradients aligned with the
anterior-posterior and medial-lateral axes.  \mpfour{These gradients may
  extend well beyond the boundary of the barrel field (e.g. as ephrin-5 does;
  see \citealp{miller_epha7-ephrin-a5_2006}) but they are shown only within
  the boundary in Fig.\,\ref{fig:main}A, to indicate that external values are
  not used by the simulation. Simulations were stepped through 30000
  iterations of Eqs.\,\ref{eq:dc}--\ref{eq:comp} ($\delta t=0.0001$)}

From initial conditions sample from a uniform random distribution for
$a(\mb{x},0)\in(0.2,0.4)$ and $c(\mb{x},0)=0$, and across a wide range of
parameter values,
%From random initial conditions for $a(\mb{x},0)$, \MPtwo{with each hexagonal element
%  of $a_i$ set from a uniform distribution according to $0.2+{U}(0,0.2)$, and
%  $c(\mb{x},0)=0$},
a clear Voronoi-like tesselation \cmnt{of topographically organised
  projections} emerged, \cmnt{confirming that barrel maps can self-organise in
  the absence of pre-specified center points. Parameters for the example
  simulation shown in Fig.\,\ref{fig:main}C (see Movie S1) were obtained by
  conducting a full parameter sweep and choosing a combination of values that
  scored well against three measures.}

% New section to describe and introduce the metrics
\metrics{First, we used an algorithm introduced by Honda to measure the
  discrepancy of each barrel shape from a Dirichletform shape. Low
  overall values of the \emph{Honda-}$\delta$ metric obtained from simulated
  barrels indicate a close correspondence of the simulated barrel field with a
  Voronoi tesselation, and thus with a biological barrel field. For the
  % Run ./build/sim/honda boundaries/rat_barrels/wb_110405_Dirichlet.svg to
  % get this result:
  tesselation overlaid on the real barrel field in Fig.\,1A, $\delta=0.025$,
  and a reduction in the Honda-$\delta$ metric in simulation
  confirmed that an equivalent `good' Voronoi pattern} (see
    \citealp{senft_mouse_1991}) \metrics{emerged within $\approx$~20000
  iterations (Fig.\,\ref{fig:main}D, red circles). Second, we devised a
  \emph{pattern difference} measure that is sensitive to deviations in the
  component shapes and overall topographic registration of two tesselations,
  $\eta$, and we used this measure to compare the simulated barrel fields to
  the real barrel field from which the simulation boundary shape was obtained
  (see Materials and Methods for details). A similar reduction in pattern
  difference over simulation time confirmed that the shapes and arrangement of
  emergent connection fields matched those of the real barrel field by around
  20000 iterations. Third, we measured the \emph{connection selectivity} at
  each location on the cortical sheet, as the connection density of the most
  dense projection divided by the sum over all projection densities, $\omega$.}
%
\metrics{Against these three metrics we are also able to characterise the
  robustness of self-organization to the model parameters, and to investigate
  the sensitivity of the model to variation in its inputs.}

% Most of this commented out region is covered in the para above.
%
%A reduction in the Honda delta
%metric (see \citealp{senft_mouse_1991}) confirmed that a `good' Voronoi
%pattern emerged within \cmnt{$\approx$~20000} iterations
%(Fig.\,\ref{fig:main}D).  A measure of the difference in \cmnt{barrel area}
%between real and simulated barrels revealed a strong correspondence (see
%Fig.\,\ref{fig:main}D, squares). \cmnt{In a chosen region, the ratio of
%  \emph{the connection density of the TC type $i$ which makes the densest
%    connections} to \emph{the sum of all connection densities} gives a measure
%  of the selectivity of the connection density, $\mu(\mb{x},t)$. We summed this measure
%  across the entire barrel domain and named it $\omega(t)$ (see
%  Fig.\,\ref{fig:main}D, grey hexagons). $\omega$ increases during the first
%  20000 simulation steps and converges on a stable
%  value. Fig.\,\ref{fig:main}E shows the selectivity measure plotted as a
%  luminance map, giving a more graded impression of the structure which
%  develops in the system.}

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig2.png}
    \caption{\MPthreePar{\textbf{A} Exploring the parameter space of the
        model given in Eqs.\,\ref{eq:dc}--\ref{eq:comp}. Graphs show the three
        pattern metrics for $t=30000$ steps as colourmaps: Honda $\delta$ (top
        row) shows conformance to a Voronoi pattern; $\eta$ (middle row) gives
        the simulated/experimental barrel area difference and $\omega$ (bottom
        row) is a measure of selectivity of the connection densities for the
        41 thalamocortical axon types. The colour maps are chosen so that for
        each row, lighter (orange and yellow) colours indicate desirable
        properties for the simulations: Lower values of $\delta$ are yellow,
        indicating good Voroni patterns; Lower values of $\eta$ indicate good
        pattern matching; Higher values of $\omega$ indicate greater
        selectivity. White squares indicate that the simulation was not stable
        for that parameter set.
%
        The parameter space explored is three dimensional with the parameter
        ratio $\alpha/\beta$ varied along the columns of graphs, the
        competition parameter $\epsilon$ varied on the $y$-axes and the
        diffusion constant, $D$ varied along the $x$-axes. An asterisk (*)
        marks the parameter set displayed in Fig.\,\ref{fig:main}. Boxes i),
        ii)~and iii)~mark parameter sets for which corresponding patterns are
        shown in \textbf{B} (for $t=30000$ steps).
%
        \textbf{B} The effect of varying the diffusion constant,
        $D$. i)~Increasing $D$ above the optimum value gives an expansion of
        the outer barrels and a corresponding compression of the barrels
        within. ii)~Conversely, an expansion of the central barrels and
        compression of those near the boundary results from reducing $D$ from
        its optimum. iii) Reducing the diffusion constant is equivalent to
        increasing the size of the domain $S$ and so this simulation, in which
        $D$ had the lowest stable value, can be thought to represent an animal
        with a large cortical area but with unchanged cortical complexity. A
        large proportion of the domain is occupied by the projections with
        intermediate interaction parameters; those with strong interaction
        parameters are compressed around the edge of the domain. A barrel
        pattern fails to form.}}
    \label{fig:paramsweep}
  \end{fullwidth}
\end{figure}

\MPthreePar{To determine the robustness of the model to its parameters, we
  investigated its behaviour with respect to changes in the parameters $D$,
  $\epsilon$ and the ratio of $\alpha$ to $\beta$. Fig.\,\ref{fig:paramsweep}A
  shows the value of the metrics $\delta$, $\eta$ and $\omega$ after 30000
  iterations, plotted as colour maps, for 216 parameter sets. The maps
  indicate that the pattern is not strongly influenced by the choice of
  $\alpha/\beta$, which was varied by 5 orders of magnitude. The most
  Voronoi-like patterns were achieved for parameter sets for which $D$ and
  $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
  preference for $D\approx0.5$, although this does not give the solution of
  maximum selectivity, which favours lower values of $D$. Thus, there is a
  trade-off between good selection and good Voroni pattern conformation.  The
  parameter set indicated by the asterisk in Fig.\,\ref{fig:paramsweep}A was
  chosen for Fig.\,\ref{fig:main} because it provides good pattern
  correspondence and a close approximation to a Voronoi tesselation.}

\MPthreePar8u{In Fig.\,\ref{fig:paramsweep}B, we present some patterns
  representing alternative parameter choices. Because the patterns are robust
  to alterations in $\epsilon$ and the choices of $\alpha$ and $\beta$, we
  concentrate our attention on the diffusion constant, $D$. Changing $D$,
  whose value defines the rate at which a region of localized branching will
  spread out across the domain, $S$, is equivalent to making changes to the
  size of $S$ with a fixed branching diffusion rate. These simulations can
  therefore be thought of as representing animals with a varying cortical
  areas but with unchanged cortical complexity. The most extreme example
  (Fig.\,\ref{fig:paramsweep}B iii) could represent an animal with a very
  large cortical area (such as the South American capybara) which receives
  sensory input from $N=41$ thalamic barreloids. Here, a large proportion of
  the domain is occupied by the projections with intermediate interaction
  parameters and those with strong interaction parameters are bunched up
  towards the edge of the domain. In this case, a barrel pattern fails to
  form, agreeing with the observation that cortical whisker barrels are absent
  in the capybara.}

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig3.png}
    \caption{
      \MPthreeSens{\textbf{A} Sensitivity of the system pattern
        metrics $\delta(t)$ (top row), $\omega(t)$ (middle row) and $\eta(t)$
        (bottom row), evaluated at $t=30000$ steps,
        to three classes of perturbation. $y$-axes and colour scales have identical ranges
        to the colour scales in Fig.\,\ref{fig:paramsweep}, to ease comparison
        with the parameter exploration.
        %
        \emph{Column 1:} Random noise drawn from a uniform distribution,
        $(\gamma_{max}-\gamma_{min})\,U(0,\nu_\gamma)$, added
        to the values for the interaction parameters, $\gamma_{i,j}$ used in
        Fig.\,\ref{fig:main}.
        %
        \emph{Column 2:} Random noise $(\rho_{max} - \rho_{min})\,U(0, \nu_{\rho})$
        was added to each hex element of $\rho_1(\mb{x})$ and $\rho_2(\mb{x})$ and the result
        was smoothed by convolution with a symmetric 2D
        Gaussian kernel of width $\sigma_\rho$. Colour maps in this column
        thus show the effect of changing the magnitude and the length scale of
        noise in the guidance fields.
        %
        \emph{Column 3:} Colour maps show the result of setting the rotational angle
        of the linearly varying guidance field, $\rho_1(\mb{x})$, to $\phi_{\rho_1}$,
        and modifying its overall gain to $G_{\rho_1}$, whilst keeping the
        parameters of $\rho_2(\mb{x})$
        unchanged from those used in the simulation in
        Fig.\,\ref{fig:main}, for which
        $\phi_{\rho_2}=84$\textdegree and $G_{\rho_2}=1$.
        %
        \textbf{B} Four ways in which the perturbations in \textbf{A} affect
        the patterns. i) The effect of significant interaction parameter noise
        causes barrels to become swapped from the positions they would
        otherwise take. ii) High magnitude, short lengthscale noise in
        $\rho(\mb{x})$ leads to `wiggly' boundaries between barrels. iii)
        Shows the effect of reducing the slope of $\rho_1$ (by a factor of 10)
        which reduces the strength of the axons' interaction with the guidance
        field in the left-right sense. Barrel rows B, C and D become `crushed'
        down the center line, and edge barrels dominate. iv) The effect of
        rotating $\rho_1$ by 20\textdegree~is a relatively mild distortion of
        the pattern in the anticlockwise sense.
      }
    }
    \label{fig:sens}
  \end{fullwidth}
\end{figure}

% This section needs to emphasize that the absolute values of gamma are not important
\MPthreeSens{We performed a sensitivity analysis to determine if the quality
  of the pattern (i.e.~the values of $\delta$, $\eta$ \& $\omega$ after
  $t=30000$ steps) was affected by perturbations to any of the following
  inputs:
%
  i)~the magnitude and offset of the noise applied to $a_i$ at $t=0$;
%
  ii)~random noise applied to the interaction parameters, $\gamma_{i,j}$;
%
  iii)~random noise (at various length scales) applied to the guidance fields;
%
  and
%
  iv)~The angle and magnitude of one of the guidance fields, ($\rho_1$;
  Fig.\,\ref{fig:main}A).
}

\MPthreeSens{First, we conducted 10 runs of the simulation shown in
  Fig.\,\ref{fig:main}C to evaluate the mean and standard deviation of the
  metrics in the default model. The results (where deviations occur as a
  result of the different random initial conditions for $a_i$) were
  $\delta=0.089\,(0.004)$, $\eta=0.2108\,(0.002)$ and
  $\omega=0.2165\,(0.0001)$. The pattern was barely affected by the random
  changes to $a_i(t=0)$ on each run.}
%
\MPthreeSens{Next, we modified the parameters of the generating uniform
  distribution to initialize $a_i$ for each hex to $0.1+{U}(0,0.4)$,
  increasing the variability of $a_i(\mb{x},t=0)$ without changing the mean
  initial activation. We ran this system 10 times and obtained the results
  $\delta=0.090\,(0.005)$, $\eta=0.2136\,(0.004)$ and
  $\omega=0.2166\,(0.0003)$. Paired two-sample t-tests demonstrated that the
  increase in noise amplitude made no significant difference to the metric
  results ($\delta: \mathrm{t}=-0.69,\mathrm{p}=0.49$, $\eta:
  \mathrm{t}=-2.0,\mathrm{p}=0.06$, $\omega:
  \mathrm{t}=-1.7,\mathrm{p}=0.11$). 10 runs of a noise-free initial condition
  with $a_i=0.3$ for every hex showed that the model is deterministic (all
  runs produced the results $\delta=0.0928$, $\eta=0.21236$ and
  $\omega=0.21649$) and that stochasticity in $a_i(\mb{x},t=0)$ is not
  required by this model, as all three results lay within one standard
  deviation of those for the runs in which $a_i$ was initialized
  stochastically.}

% Quick check that noise free result is within 1 SD for the noisier initial conditions
%         Noisefree    Normal noise    NF SD from Normal?   Noisy            NF SD from Noisy?
% delta   0.0928       0.089(0.004)    <1                   0.090 (0.005)    <1
% eta     0.2124       0.2108(0.002)   <1                   0.2136 (0.004)   <1
% omega   0.21649      0.2165(0.0001)  <1                   0.2166 (0.0003)  <1

\MPthreeSens{Fig\,\ref{fig:sens}A shows the results for perturbations ii) to
  iv). The results show that the Voronoi nature of the pattern ($\delta$) and
  the selectivity of the fields are not strongly affected by interaction
  parameter noise, but the pattern metric, $\eta$ shows a degradation in the
  pattern arrangement when a) noise in the interaction parameters causes their
  disordering (find better way to express?), b) noise in the guidance field
  with a high magnitude and short length scale causes disruption to the barrel
  edges and c) the guidance fields tend towards a non-orthogonal arrangement,
  or if one field becomes substantially stronger in its interaction with the
  thalamocortical projections. Fig\,\ref{fig:sens}B shows perturbed examples.}

To further investigate the interplay of genetic and intrinsic mechanisms we
simulated a seminal barrel duplication experiment
\citep{shimogori_fibroblast_2005}, first by solving on a cortical domain
comprising two separate, identically shaped cortical boundaries. In this case
\cmnt{NOTE: The following only holds for the divisive normalization system -
  in the comp2 system, there's no `coupling' at all between the two regions,
  so the mirrored region arranges itself as expected with the one reversed
  guidance field.} one distinct cortical field per thalamic projection emerged, within one of the
two boundaries only, i.e., no duplication. However, merging the two fields to
create an extended mirror-symmetric boundary shape, with an anterior-posterior
guidance field reflected at the join, gave rise to two mirror-symmetrical
barrel fields comprising 2N barrels, i.e., two cortical fields for each
thalamic projection (Fig.\,\ref{fig:fgf8}). Together these results suggest that by the
misexpression of Fgf8, \cite{shimogori_fibroblast_2005}
effectively created one large barrel field rather than two distinct cortical
fields. \cmnt{Add something about the slightly off-vertical arrangement of the
  fields.}

% This figure is text width
\begin{center}
  \includegraphics[width=\linewidth]{./Fig4.png}
  \captionof{figure}{Guidance
    fields and emergent barrel pattern in a Fgf8 misexpression experiment
    (c.f. \citealp{assimacopoulos_fibroblast_2012}), simulated by reflecting
    $\rho_1$ from Fig.\,\ref{fig:main}A at the join. \cmnt{The model
      parameters match those in Fig.\,\ref{fig:main}C. Scale
      bars are 1\,mm.}}
  \label{fig:fgf8}
\end{center}

Add paragraph on whisker trimming here.

\begin{center}
  \includegraphics[width=\linewidth]{./Fig5.png}
  \captionof{figure}{\MPtwo{Simulating whisker trimming by reduction of
      $\epsilon$ for the thalamocortical projection for barrel C3, for which
      $i=13$. All other model parameters match those in
      Fig.\,\ref{fig:main}C. \textbf{A} Graph of individual barrel measures
      plotted vs.~the multiplier applied to $\epsilon_{13}$. \textbf{B} The
      pattern for $\epsilon_{m}=0.64$.}}
  \label{fig:whiskertrim}
\end{center}

\section{Discussion}

The present results suggest that the key requirements for the emergence of
realistic barrel patterning are i)~at each cortical location thalamocortical
projections compete for a limited number of available synaptic connections
(Eqs.~\ref{eq:dc}--\ref{eq:da}), ii)~at each location the branching rate of a
given projection is reduced by the density of other projections
(Eq.\,\ref{eq:comp}), and iii)~the branch density of each projection is
conserved over time.

The emergence of barrels in simulation required competition between thalamic
projections in terms of synaptic connectivity and also competition in terms of
cortical space, as represented by $\chi$, with an implicit requirement for a
self/other identifier amoungst projections. This latter form of competition may account for the absence of barrels in rodents with larger brains, such as
capybara, for which competition for space is presumably weaker
\citep{woolsey_comparative_1975}. Hence, irrespective of whether barrels are
necessary for adaptive whisker function, the emergence of somatotopically
ordered modular structures may be an inevitable consequence of local
competition for cortical territory driven by input from an array of discrete
sensory organs \citep{purves_iterated_1992}.

\cmnt{In reality, scores of smaller Dirichletform barrels representing the
  microvibrissae form to the right of where the E-row barrels emerge in our
  simulations, presumably via the same competitive processes. Enforcing here
  the same boundary condition as used to represent the true edges of the
  barrel field is necessary to ensure the stability of the simulation, though
  we acknowledge that this region of the boundary is enforced primarily to
  keep the number of simulated projections, and hence the overall
  computational complexity of the simulation, manageable (simulating each
  additional projection introduces an extra \textbf{YYY} new dynamical
  variables).}

It is important to emphasize that the formulation of the model is entirely
local, insofar as simulation requires no information to be communicated from a
given cortical grid cell to any but those immediately adjacent (via
diffusion). Hence the simulations demonstrate how a self-organizing system,
constrained by genetically specified guidance cues and by the shape of the
cortical field boundary, can faithfully reproduce an arrangement of cell
aggregates in one neural structure as a topographic map in another.

Moreover, the present results confirm that somatotopic map formation does not
require the pre-specification of center-points by as yet undetermined
additional developmental mechanisms.

\section{Materials \& Methods}

% Be sure to include/mention:
%
% The no-flux boundary conditions applied in the \cite{karbowski_model_2004}
% model are retained.
%
% Arbitrary boundary shapes were applied

The cortical sheet was modelled as a two dimensional hexagonal lattice, which
simplifies the computation of the 2D Laplacian. Within a boundary traced
around the edge of a
rat barrel field (Fig.\,\ref{fig:main}A) we set the hex-to-hex distance
$d$ to 0.03\,mm, which resulted in a lattice containing 6515 hexes for the
simulations shown in Figs.\,\ref{fig:main}A,C \& D and 12739 hexes for the Fgf8
misexpression study shown in Fig.\,\ref{fig:fgf8}. Each hex contained 82 time-dependent
variables: 41 branching densities ($a_i$) and 41 connection densities ($c_i$).
The rate of change of each of the time-dependent variables (Eqs.\,\ref{eq:dc}
\& \ref{eq:da}) was computed using a fourth-order Runge-Kutta method.

The most involved part of this computation is to find the divergence of the
flux of axonal branching, $\mb{J}_i(\mb{x},t)$, the term in parentheses in
Eq.\,\ref{eq:da}:
%
\begin{equation}
  \label{eq:divJ}
  \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla
  a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \color{revblue}+
  \frac{\epsilon a_i}{N-1} \nabla \hat{a}_i \color{black}\right),
\end{equation}
%
\MPtwo{where $\hat{a}_i\equiv\sum_{j\ne i}^{N}a_j$.} Note that the sum of the
guidance gradients is time-independent and define $\mb{g}_i(\mb{x}) \equiv
\sum_{j=1}^{M} \gamma_{i,j} \nabla\rho_j(\mb{x})$.  Because the divergence
operator is distributive, Eq.\,\ref{eq:divJ} can be expanded using vector
calculus identities (dropping references to $\mb{x}$ and $t$ for clarity):
%
\begin{equation}
\nabla\vcdot\mb{J}_i = \nabla\vcdot\big(D \nabla a_i\big) -
\nabla\vcdot\big(a_i \mb{g}_i\big) \color{revblue} + \frac{\epsilon}{N-1}\nabla\vcdot\big({a}_i\nabla\hat{a}_i\big)\color{black}.
\end{equation}
%
Applying the vector calculus product rule identity yields
%
\begin{equation} \label{eq:divJExpanded}
%
\nabla\vcdot\mb{J}_i =
%
D \dvrg a_i % term1
-
a_i\nabla\vcdot\mb{g}_i % term2
-
\mb{g}_i\vcdot\nabla a_i % term3
\color{revblue}
+ \frac{\epsilon a_i}{N-1} \dvrg \hat{a}_i % term1_1
+ \frac{\epsilon}{N-1} \nabla \hat{a}_i\vcdot\nabla{a}_i% term1_2
\color{black}
,
\end{equation}
%
which has \MPtwo{five} elements to compute: i)~$D \dvrg a_i$ (the Laplacian of
$a_i$); ii)~a time-independent modulator of $a_i$ (because
$\nabla\vcdot\mb{g}_i$ is a time-independent static field); iii)~the scalar
product of the static vector field $\mb{g}_i$ and the gradient of $a_i$;
\MPtwo{iv)~the Laplacian of $\hat{a}_i$; and v)~a term involving the gradients
  of $a_i$ and $\hat{a}_i$}. Each of the divergences can be simplified by
means of Gauss's Theorem following \cite{lee_hexagonal_2014}.

(i) The computation of the mean value of the Laplacian across one hexagon of
area $\Omega = \frac{\sqrt{3}}{2}d^2$, located at position $\mb{p}_0$, with
neighbours at positions $\mb{p}_1$--$\mb{p}_6$ is
%
\begin{equation} \label{eq:lapl}
\begin{split}
\langle D \dvrg a_i(\mb{p}_0,t) \rangle  = \frac{D}{\Omega} \oiint_\Omega \dvrg a_i(\mb{x}, t) \dif\Omega & = \frac{D}{\Omega} \oint \frac{\partial a_i}{\partial \hat{\mb{n}}} \dif\gamma \\
& \approx \frac{D}{\Omega} \sum_{j=1}^6 \frac{\partial a_i(\mb{p}_j)}{\partial \hat{\mb{n}}} \bigg\rvert_{\mathrm{mid}} v \\
& = \frac{2D}{\sqrt{3} d^2} \sum_{j=1}^6 \frac{a_i(\mb{p}_j) - a_i(\mb{p}_0)}{d} \frac{d}{\sqrt{3}} \\
& = \frac{2D}{3 d^2} \sum_{j=1}^6 \big(a_i(\mb{p}_j) - a_i(\mb{p}_0)\big),
\end{split}
\end{equation}
%
where $v = d/\sqrt{3}$ is the length of each edge of the hexagon and $\dif\gamma$
is an infinitesimally small distance along its perimeter.

\noindent
ii) The computation of the second term in Eq.\,\ref{eq:divJExpanded},
$\langle a_i(\mb{p}_0,t)\nabla\vcdot\mb{g}_i(\mb{p}_0)\rangle$, can be written out similarly:

\begin{equation} \label{eq:divg}
\begin{split}
%
\frac{1}{\Omega} \oiint_\Omega a_i\nabla\vcdot\mb{g}_i\;\dif\Omega & = \frac{a_i(\mb{p}_0,t)}{\Omega}  \oint \mb{g}_i\vcdot \dif\hat{\mathbf{n}} \\
%
& \approx \frac{a_i(\mb{p}_0,t)}{\Omega} \sum_{j=1}^{6} \frac{\mb{g}_i(\mb{p}_j) + \mb{g}_i(\mb{p}_0)}{2}\vcdot \hat{\mb{n}}\;v \\
%
& = \frac{2 a_i(\mb{p}_0,t)v}{\sqrt{3}d^2} \sum_{j=1}^{6} \bigg[ \frac{g_i^x(\mb{p}_j) + g_i^x(\mb{p}_0)}{2} \vcdot  \hat{\mb{n}} + \frac{g_i^y(\mb{p}_j) + g_i^y(\mb{p}_0)}{2} \vcdot  \hat{\mb{n}} \bigg] \\
%
%%& = \frac{2 a_i(\mb{p}_0,t)d}{\sqrt{3}d^2\sqrt{3}} \bigg( \sum_{j=1}^{6} \frac{\mb{g}_j^x + \mb{g}_0^x}{2} \vcdot  \hat{\mb{n}} + \sum_{j=1}^{6} \frac{\mb{g}_j^y + \mb{g}_0^y}{2} \vcdot  \hat{\mb{n}} \bigg) \\
%
%\Rightarrow \langle a_i(\mb{p}_0,t)\nabla\vcdot\mb{g}(\mb{p}_0)\rangle &
%\approx \frac{a_i(\mb{p}_0,t)}{3d} \bigg( \sum_{j=1}^{6} \big(g_i^x(\mb{p}_j)
%+ g_i^x(\mb{p}_0)\big) \cos \big(\frac{\pi}{3}(j-1)\big) + \sum_{j=1}^{6}
%\big({g_i^y(\mb{p}_j) + g_i^y(\mb{p}_0)}\big) \sin
%\big(\frac{\pi}{3}(j-1)\big) \bigg), \\
%% Condensing into a single sum:
\Rightarrow \langle a_i(\mb{p}_0,t)\nabla\vcdot\mb{g}(\mb{p}_0)\rangle & \approx \frac{a_i(\mb{p}_0,t)}{3d} \sum_{j=1}^{6} \bigg[ \big(g_i^x(\mb{p}_j) + g_i^x(\mb{p}_0)\big) \cos \big(\frac{\pi}{3}(j-1)\big) + \big({g_i^y(\mb{p}_j) + g_i^y(\mb{p}_0)}\big) \sin \big(\frac{\pi}{3}(j-1)\big) \bigg], \\
%
\end{split}
\end{equation}
%
where $g_i^x$ and $g_i^y$ are the Cartesian components of $\mb{g}_i$. Both this
last expression, and the final expression of Eq.\,\ref{eq:lapl} can
be computed locally, by summing over values of the nearest neighbours.

\noindent
iii) The \MPtwo{middle} term in Eq.\,\ref{eq:divJExpanded} is the scalar product of two
vector fields which is straightforward to compute from their Cartesian
components.

\noindent
\MPtwo{iv) The same method used to compute $\dvrg a_i$ in term (i) is used to compute
$\dvrg \hat{a}_i$.}

\noindent
\MPtwo{v) The final term is the scalar product of the two vector fields $\nabla
  a_i$ and $\nabla \hat{a}_i$.}

By separating the computation of Eq.\,\ref{eq:divJ} into parts (i)--(v), the
no-flux boundary condition,
%
\begin{equation}
  \label{eq:noflux}
  \mb{J}_i(\mb{x},t)\big\rvert_{\mathrm{boundary}} = 0,
\end{equation}
%
can be fulfilled. On
the boundary, the contribution to $\mb{J}$ resulting from the first term of
Eq.\,\ref{eq:divJExpanded} can be fixed to 0 by the `ghost cell method' in
which, during the evaluation of (i), a hex outside the boundary containing the
same value as the hex inside the boundary is imagined to exist such that the
flux of $\mb{J}$ across the boundary is 0. Then, $\mb{g}_i(\mb{x})$ can be
tailored so that it, and its normal derivative approach 0 at the boundary,
ensuring that the second and third terms of Eq.\,\ref{eq:divJExpanded} also
contribute nothing to $\mb{J}$. This is achieved by \MPtwo{multiplying}
$\mb{g}_i(\mb{x})$ \MPtwo{by} a sharp logistic function of the distance, $d_b$,
from $\mb{x}$ to the boundary, \MPtwo{of the form $1/[1 +
  \exp(100(d_f-d_b))]$, where $d_f=0.1$\,mm is the boundary fall-off distance,
  chosen to be approximately $3d$}.

\metrics{The pattern-matching metric, $\eta$, incorporates information about i)
  the differences in the areas of the simulated and experimental barrels
  ($\mathcal{A}_i^{\mathrm{sim}}$ and $\mathcal{A}_i^{\mathrm{exp}}$); and ii)
  the information provided by `adjacency vectors', $\mathcal{V}_i$ (one for
  each barrel) whose $N$ elements contain the length of the barrel boundary
  which adjoins each other barrel in the pattern. The value of
  $\mathcal{V}_{i,j}$ (the $j^{\mathrm{th}}$ element of $\mathcal{V}_i$) is
  the length of the boundaries between regions of type $i$ and regions of type
  $j$. In a well-formed barrel pattern, $\mathcal{V}_{i}$ is a sparse vector
  and note that $\mathcal{V}_{i,i} = 0, \forall i$. $\mathcal{V}_{i}$ can be
  computed for the experimental and simulated barrel patterns and compared. A
  dimensionless measure of the pattern's arrangement and neighbour relations
  can be obtained from the scalar product of the simulated and experimental
  adjacency vectors:
  $\frac{1}{N} \sum_i \frac { \mathcal{V}_{i}^{\mathrm{sim}}}
  {b_{i}^{\mathrm{sim}} } \vcdot \frac{\mathcal{V}_{i}^{\mathrm{exp}}}{b_{i}^{\mathrm{exp}}}$, where
  $b_i^{\mathrm{exp}}$($b_i^{\mathrm{sim}}$) is the total length of the boundary of
  experimental(simulated) barrel $i$. This measure is low early in the simulation, when the regions do
  not have the neighbour relations of the experimental pattern
  (Fig.\,\ref{fig:main}C, $t=1000$) and maximal when the neighbour relations
  have become established ($t=10000$). It does not distinguish between
  $t=10000$ and $t=30000$.}
%
\metrics{A second way to analyse the adjacency vectors is to consider
  the mean magnitude of the difference between the simulated and experimental
  vectors: $\frac{1}{N} \sum_i \norm{ \mathcal{V}_{i}^{\mathrm{sim}} -
    \mathcal{V}_{i}^{\mathrm{exp}} } $. This tends to 0 for a perfect
  match and can separate patterns with straight boundaries from those with
  noise on the boundary shape (For an example of `noisy edges', see Fig.\,\ref{fig:sens}B.ii).}
%
\metrics{We combined all of these measures into a single pattern quality metric:}
%
\begin{equation} \label{eq:eta}
\color{revred}
  \eta =
  \frac
  {
    \frac{1}{N} \sum_i \norm[\big]{ \mathcal{A}_i^{\mathrm{sim}} - \mathcal{A}_i^{\mathrm{exp}} }
    \times
    \frac{1}{N} \sum_i \norm[\big]{ \mathcal{V}_{i}^{\mathrm{sim}} -
      \mathcal{V}_{i}^{\mathrm{exp}} }
  }
  {
    \frac{1}{N}
    \sum_i \frac {\mathcal{V}_{i}^{\mathrm{sim}}} {b_{i}^{\mathrm{sim}}}
    \vcdot
    \frac {\mathcal{V}_{i}^{\mathrm{exp}}} {b_{i}^{\mathrm{exp}}}
  },
\color{black}
\end{equation}
%
\metrics{which has units of mm$^3$.}

All code required to reproduce these results is available at
\url{https://github.com/ABRG-Models/BarrelEmerge/tree/eLife}. The computations
described in (i)--(v) may be found in the class method
\code{RD\_James\MPtwo{\_comp2}::compute\_divJ()} which calculates \code{term1},
\code{term2}, \code{term3}, \code{term1\_1} and \code{term1\_2}, respectively.

\section{Movie S1 caption}

Movie corresponding to Fig.\,\ref{fig:main}C in the main paper. Simulation parameters were
$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
$\epsilon=1.2$ and $\delta{t}=0.0001$. Colours indicate the thalamic
projection for which the connection density is maximal, black lines delineate
boundaries, and overlaid contours show $c>0.5$. The final frame in the movie
is step 30000 of the simulation.

\section{Acknowledgments}

The authors thank Jason Berwick at the University of Sheffield for advice and
for access to the rat barrel stains used to construct Fig.\,1A. This work was
supported by a Collaborative Activity Award, \emph{Cortical Plasticity Within
  and Across Lifetimes}, from the James S.~McDonnell Foundation (grant
220020516).

\bibliography{../BarrelEmerge}

\end{document}

% NOTES

%\MPtwo{In other words, we simply scaled the barreloid coordinates in an anterior-posterior/medial-lateral coordinate system to create the $\gamma$ values in the range [-2,2]. Note that positive \emph{and} negative $\gamma$ values were generated; the origin of the coordinate system was located at the approximate center of the thalamic barreloid field. This prevented an overall drift of the cortical pattern away from the centroid of the barrel field and may explain why coordinate-giving transcription factors and the ephrins are often found as pairs acting in opposing directions.}  \MPtwo{We concentrated on macrovibrissae (excluding microvibrissae) because it is easy to identify their corresponding structures both in the thalamus and in the cortex. We also excluded the 5 rhinal whiskers because their cortical representation is in a separate island, distinct, and somewhat removed from the representation of the macrovibirissae.}}

%\MPtwo{Thus, we suggest that there exist a pair of signalling molecules whose expression varies linearly (with orthogonal gradients) across the thalamus, analgous to the ephrins, whose existence is well known in the cortex. Thalamocortical axons growing from a region in the thalamus with high expression of one of these molecules will interact with a corresponding cortical signalling molecule more strongly than an axon emanating from a region with low expression of the thalamic signal.}

%\MPtwo{We defined two orthogonal, linear gradients oriented in the anterior-posterior and medial-lateral directions and used the coordinates of 41 barreloid centers recreated from \cite{haidarliu_size_2001}, corresponding to the macrovibrissae, to populate 41 pairs of $\gamma$ values.}  (Fig.\,\ref{fig:main}B). \MPtwo{In other words, we simply scaled the barreloid coordinates in an anterior-posterior/medial-lateral coordinate system to create the $\gamma$ values in the range [-2,2]. Note that positive \emph{and} negative $\gamma$ values were generated; the origin of the coordinate system was located at the approximate center of the thalamic barreloid field. This prevented an overall drift of the cortical pattern away from the centroid of the barrel field and may explain why coordinate-giving transcription factors and the ephrins are often found as pairs acting in opposing directions.}  \MPtwo{We concentrated on macrovibrissae (excluding microvibrissae) because it is easy to identify their corresponding structures both in the thalamus and in the cortex. We also excluded the 5 rhinal whiskers because their cortical representation is in a separate island, distinct, and somewhat removed from the representation of the macrovibirissae.}